---
title: "*R Code for Scientific Article*"
subtitle: '*Bayesian Hierarchical Model for Immune Responses to Leishmania - a Tick Borne Coinfection Study'
author: '**Felix Pabon-Rodriguez**'
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: no
  pdf_document:
    toc: no
header-includes:
- \usepackage{amsmath}
- \usepackage{geometry}
- \usepackage{xcolor}
- \usepackage{booktabs}
- \usepackage{float}
- \usepackage{bbm}
geometry: margin=1in
---

```{r Setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.pos = "H")

# Loading libraries
library(readxl)
library(dplyr)
library(coda) 
library(rjags)
library(MASS)
library(ggplot2)
library(R2jags)
library(lattice)
library(bayesplot)
library(BayesPostEst)
library(ggmcmc)
library(RCurl)
library(truncnorm)
library(kableExtra)
library(mvtnorm)
library(rlist)
library(extraDistr)
library(msm)
library(tmvtnorm)
library(runjags)
library(plotrix)
library(MCMCvis)

# use deparse and substitute to get the name of a function argument
getname <- function(xobj) {deparse(substitute(xobj))}

```

\ 

# Descriptive Summary

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Data set
leish <- read.csv(file = "./Data/Sarolaner Cohort Data with missing values.csv", 
                  header = TRUE, 
                  na.strings = "", 
                  stringsAsFactors = TRUE)

# Dogs with ID 27, 91, 737, 743, 750, 758, 761, 770, 773 lost to follow-up
#leish <- leish[!(leish$ID %in% c(737, 750, 758, 761, 770, 773)),]

# New variable: disease_status per time point, based on leishvet score
# healthy = 1 if leishvet score is 0 or 1
# asymptomatic (inflammatory) = 2 if leishvet score is 1 or 2  
# symptomatic (regulatory) = 3 if leishvet score is 3 or 4  
leish$DStatus_TP1 <- ifelse(leish$LeishVet.Score.TP1 %in% c(0,1), 1, 
                            ifelse(leish$LeishVet.Score.TP1 == 2, 2, 3))
leish$DStatus_TP3 <- ifelse(leish$LeishVet.Score.TP3 %in% c(0,1), 1, 
                            ifelse(leish$LeishVet.Score.TP3 == 2, 2, 3))
leish$DStatus_TP4 <- ifelse(leish$LeishVet.Score.TP4 %in% c(0,1), 1, 
                            ifelse(leish$LeishVet.Score.TP4 == 2, 2, 3))
leish$DStatus_TP5 <- ifelse(leish$LeishVet.Score.TP5 %in% c(0,1), 1, 
                            ifelse(leish$LeishVet.Score.TP5 == 2, 2, 3))
leish$DStatus_TP6 <- ifelse(leish$LeishVet.Score.TP6 %in% c(0,1), 1, 
                            ifelse(leish$LeishVet.Score.TP6 == 2, 2, 3))
leish$DStatus_TP7 <- ifelse(leish$LeishVet.Score.TP7 %in% c(0,1), 1, 
                            ifelse(leish$LeishVet.Score.TP7 == 2, 2, 3))

# Add some unobserved variables
leish$DStatus_TP2 <- rep(NA,nrow(leish)) 

# Add removed state (Disease Status = 4, if dog is removed due to severe leish)
# dog IDs: 758, 761, 743
leish$DStatus_TP7[leish$ID == "758"] = 4
leish$DStatus_TP7[leish$ID == "743"] = 4
leish$DStatus_TP5[leish$ID == "761"] = 4
leish$DStatus_TP6[leish$ID == "761"] = 4
leish$DStatus_TP7[leish$ID == "761"] = 4


#write.csv(leish, file = "dataset_for_julia.csv",
#          col.names = TRUE, row.names = FALSE,
#          na = "")
```

Data:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
leish
```

\ 

(a)  Age Group of the Dogs

```{r, echo=FALSE, message=FALSE, warning=FALSE}
leish %>%
  group_by(Age_group) %>%
  summarise(n = n(), prop = (n/50)*100)
```

\ 

(b) Number of Dogs that Ever SNAP Positive (Coinfection Positive)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
leish %>%
  group_by(Ever.SNAP.Positive) %>%
  summarise(n = n(), prop = (n/50)*100)
```

\ 

(c) Types of Coinfection Identified per Time Point

```{r, echo=FALSE}
table(leish$SNAP_TP1)
table(leish$SNAP_TP2)
table(leish$SNAP_TP3)
table(leish$SNAP_TP4)
table(leish$SNAP_TP5)
table(leish$SNAP_TP6)
table(leish$SNAP_TP7)
```

\ 

(d) Mean of PCR Load per Point Time (# parasites / ML blood)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
leish %>%
  summarise_at(c("PCR_Load_TP1", "PCR_Load_TP2", 
                 "PCR_Load_TP3", "PCR_Load_TP4",
                 "PCR_Load_TP5", "PCR_Load_TP6",
                 "PCR_Load_TP7"), 
               mean, na.rm = TRUE)
```

\ 

(e) SD of PCR Load per Point Time (# parasites / ML blood)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
leish %>%
  summarise_at(c("PCR_Load_TP1", "PCR_Load_TP2", 
                 "PCR_Load_TP3", "PCR_Load_TP4",
                 "PCR_Load_TP5", "PCR_Load_TP6",
                 "PCR_Load_TP7"), 
               sd, na.rm = TRUE)
```

\ 

(f) PCR Result when Enroll (Presence of Parasitemia)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
leish %>%
  group_by(PCR_Result_Enroll) %>%
  summarise(n = n()) 
```

\ 

(g) DPP Results when Enroll (Presence of anti-Leishmania antibodies)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
leish %>%
  group_by(DPP_Enroll.Result) %>%
  summarise(n = n())
```

\  

(h) PCR Results per Point Time

```{r, echo=FALSE}
pcr_pos <- c(sum(1*(leish$PCR_Result_TP1 == "P"), na.rm = TRUE),
              sum(1*(leish$PCR_Result_TP2 == "P"), na.rm = TRUE),
              sum(1*(leish$PCR_Result_TP3 == "P"), na.rm = TRUE),
              sum(1*(leish$PCR_Result_TP4 == "P"), na.rm = TRUE),
              sum(1*(leish$PCR_Result_TP5 == "P"), na.rm = TRUE),
              sum(1*(leish$PCR_Result_TP6 == "P"), na.rm = TRUE),
              sum(1*(leish$PCR_Result_TP7 == "P"), na.rm = TRUE))
pcr_neg <- c(sum(1*(leish$PCR_Result_TP1 == "N"), na.rm = TRUE),
              sum(1*(leish$PCR_Result_TP2 == "N"), na.rm = TRUE),
              sum(1*(leish$PCR_Result_TP3 == "N"), na.rm = TRUE),
              sum(1*(leish$PCR_Result_TP4 == "N"), na.rm = TRUE),
              sum(1*(leish$PCR_Result_TP5 == "N"), na.rm = TRUE),
              sum(1*(leish$PCR_Result_TP6 == "N"), na.rm = TRUE),
              sum(1*(leish$PCR_Result_TP7 == "N"), na.rm = TRUE))
pcr_bl <- c(sum(1*(leish$PCR_Result_TP1 == "BL"), na.rm = TRUE),
              sum(1*(leish$PCR_Result_TP2 == "BL"), na.rm = TRUE),
              sum(1*(leish$PCR_Result_TP3 == "BL"), na.rm = TRUE),
              sum(1*(leish$PCR_Result_TP4 == "BL"), na.rm = TRUE),
              sum(1*(leish$PCR_Result_TP5 == "BL"), na.rm = TRUE),
              sum(1*(leish$PCR_Result_TP6 == "BL"), na.rm = TRUE),
              sum(1*(leish$PCR_Result_TP7 == "BL"), na.rm = TRUE))
data.frame(Time.Points = 1:7,
           PCR.Negative = pcr_neg,
           PCR.Positive = pcr_pos,
           PCR.Borderline.BL = pcr_bl,
           PCR.NA = nrow(leish) - pcr_neg - pcr_pos - pcr_bl) 
```

\ 

(i) LeishVet Score per Point Time

```{r, echo=FALSE}
leish_vet0 <- c(sum(1*(leish$LeishVet.Score.TP1 == 0), na.rm = TRUE),
                NA,
              sum(1*(leish$LeishVet.Score.TP3 == 0), na.rm = TRUE),
              sum(1*(leish$LeishVet.Score.TP4 == 0), na.rm = TRUE),
              sum(1*(leish$LeishVet.Score.TP5 == 0), na.rm = TRUE),
              sum(1*(leish$LeishVet.Score.TP6 == 0), na.rm = TRUE),
              sum(1*(leish$LeishVet.Score.TP7 == 0), na.rm = TRUE))

leish_vet1 <- c(sum(1*(leish$LeishVet.Score.TP1 == 1), na.rm = TRUE),
              NA,
              sum(1*(leish$LeishVet.Score.TP3 == 1), na.rm = TRUE),
              sum(1*(leish$LeishVet.Score.TP4 == 1), na.rm = TRUE),
              sum(1*(leish$LeishVet.Score.TP5 == 1), na.rm = TRUE),
              sum(1*(leish$LeishVet.Score.TP6 == 1), na.rm = TRUE),
              sum(1*(leish$LeishVet.Score.TP7 == 1), na.rm = TRUE))

leish_vet2 <- c(sum(1*(leish$LeishVet.Score.TP1 == 2), na.rm = TRUE),
              NA,
              sum(1*(leish$LeishVet.Score.TP3 == 2), na.rm = TRUE),
              sum(1*(leish$LeishVet.Score.TP4 == 2), na.rm = TRUE),
              sum(1*(leish$LeishVet.Score.TP5 == 2), na.rm = TRUE),
              sum(1*(leish$LeishVet.Score.TP6 == 2), na.rm = TRUE),
              sum(1*(leish$LeishVet.Score.TP7 == 2), na.rm = TRUE))

leish_vet3 <- c(sum(1*(leish$LeishVet.Score.TP1 == 3), na.rm = TRUE),
              NA,
              sum(1*(leish$LeishVet.Score.TP3 == 3), na.rm = TRUE),
              sum(1*(leish$LeishVet.Score.TP4 == 3), na.rm = TRUE),
              sum(1*(leish$LeishVet.Score.TP5 == 3), na.rm = TRUE),
              sum(1*(leish$LeishVet.Score.TP6 == 3), na.rm = TRUE),
              sum(1*(leish$LeishVet.Score.TP7 == 3), na.rm = TRUE))

leish_vet4 <- c(sum(1*(leish$LeishVet.Score.TP1 == 4), na.rm = TRUE),
              NA,
              sum(1*(leish$LeishVet.Score.TP3 == 4), na.rm = TRUE),
              sum(1*(leish$LeishVet.Score.TP4 == 4), na.rm = TRUE),
              sum(1*(leish$LeishVet.Score.TP5 == 4), na.rm = TRUE),
              sum(1*(leish$LeishVet.Score.TP6 == 4), na.rm = TRUE),
              sum(1*(leish$LeishVet.Score.TP7 == 4), na.rm = TRUE))

leish_vetNA <- c(sum(1*is.na(leish$LeishVet.Score.TP1)),
                 nrow(leish),
                 sum(1*is.na(leish$LeishVet.Score.TP3)),
                 sum(1*is.na(leish$LeishVet.Score.TP4)),
                 sum(1*is.na(leish$LeishVet.Score.TP5)),
                 sum(1*is.na(leish$LeishVet.Score.TP6)),
                 sum(1*is.na(leish$LeishVet.Score.TP7))) 

data.frame(TP = 1:7,
           LVet_0 = leish_vet0,
           LVet_1 = leish_vet1,
           LVet_2 = leish_vet2,
           LVet_3 = leish_vet3,
           LVet_4 = leish_vet4,
           LVet_NA = leish_vetNA) 
```

\ 

(j) Disease Status per Time Point
```{r, echo=FALSE, message=FALSE, warning=FALSE}
table(leish$DStatus_TP1, useNA = "ifany")
table(leish$DStatus_TP2, useNA = "ifany")
table(leish$DStatus_TP3, useNA = "ifany")
table(leish$DStatus_TP4, useNA = "ifany")
table(leish$DStatus_TP5, useNA = "ifany")
table(leish$DStatus_TP6, useNA = "ifany")
table(leish$DStatus_TP7, useNA = "ifany")
```

\ 

# Longitudinal Plots

```{r, echo=FALSE}
ggbg <- function(color) {
    points(0, 0, pch=16, cex=1e6, col=color)
    grid(col="white", lty=1)
}

pcr_load <- as.data.frame(with(leish,cbind(ID,PCR_Load_TP1, PCR_Load_TP2, PCR_Load_TP3,
                             PCR_Load_TP4, PCR_Load_TP5, PCR_Load_TP6,
                             PCR_Load_TP7)))
antib <- as.data.frame(with(leish, cbind(ID, SLA.ELISA.OD.ratio_TP1,SLA.ELISA.OD.ratio_TP2,
                           SLA.ELISA.OD.ratio_TP3,SLA.ELISA.OD.ratio_TP4,
                           SLA.ELISA.OD.ratio_TP5,SLA.ELISA.OD.ratio_TP6,
                           SLA.ELISA.OD.ratio_TP7)))
dstat <- as.data.frame(with(leish, cbind(ID, DStatus_TP1, DStatus_TP2, DStatus_TP3,
                           DStatus_TP4, DStatus_TP5, DStatus_TP6, DStatus_TP7)))



long_pcrload <- reshape(pcr_load, 
                         direction = "long",
                         varying = list(names(pcr_load)[2:8]),
                         v.names = "Value",
                         idvar = c("ID"),
                         timevar = "TimePoint",
                         times = 1:7)
long_pcrload <- long_pcrload[order(long_pcrload$ID),] 


long_antib <- reshape(antib, 
                      direction = "long",
                      varying = list(names(antib)[2:8]),
                      v.names = "Value",
                      idvar = c("ID"),
                      timevar = "TimePoint",
                      times = 1:7)
long_antib <- long_antib[order(long_antib$ID),] 


long_dstat <- reshape(dstat, 
                      direction = "long",
                      varying = list(names(dstat)[2:8]),
                      v.names = "Value",
                      idvar = c("ID"),
                      timevar = "TimePoint",
                      times = 1:7)
long_dstat <- long_dstat[order(long_dstat$ID),] 


mean_PCR_load <- c(mean(leish$PCR_Load_TP1, na.rm = TRUE),
                   mean(leish$PCR_Load_TP2, na.rm = TRUE),
                   mean(leish$PCR_Load_TP3, na.rm = TRUE),
                   mean(leish$PCR_Load_TP4, na.rm = TRUE),
                   mean(leish$PCR_Load_TP5, na.rm = TRUE),
                   mean(leish$PCR_Load_TP6, na.rm = TRUE),
                   mean(leish$PCR_Load_TP7, na.rm = TRUE))

mean_antibodies <- c(mean(leish$SLA.ELISA.OD.ratio_TP1, na.rm = TRUE),
                   mean(leish$SLA.ELISA.OD.ratio_TP2, na.rm = TRUE),
                   mean(leish$SLA.ELISA.OD.ratio_TP3, na.rm = TRUE),
                   mean(leish$SLA.ELISA.OD.ratio_TP4, na.rm = TRUE),
                   mean(leish$SLA.ELISA.OD.ratio_TP5, na.rm = TRUE),
                   mean(leish$SLA.ELISA.OD.ratio_TP6, na.rm = TRUE),
                   mean(leish$SLA.ELISA.OD.ratio_TP7, na.rm = TRUE))

par(mfrow=c(1,2))
interaction.plot(long_pcrload$TimePoint, long_pcrload$ID, long_pcrload$Value, 
                 xlab="Time Point", ylab="PCR Load", legend=F, lty = 1, cex = 4, 
                 lwd = 2, main = "Pathogen Load Over Time", 
                 col = rgb(255, 0, 0, max = 255, alpha = 150, names = "red"),
                 panel.first = ggbg("gray92")) 
lines(mean_PCR_load, col = "black", lwd = 3)
grid(NULL, NULL, col = "darkgray")

interaction.plot(long_antib$TimePoint, long_antib$ID, long_antib$Value, 
                 xlab="Time Point", ylab="Antibody Levels", legend=F, lty = 1, 
                 lwd = 2, main = "Antibody Levels Over Time", cex = 4,
                 col = rgb(0, 102, 0, max = 255, alpha = 150, names = "darkgreen"), 
                panel.first = ggbg("gray92")) 
lines(mean_antibodies, col = "black", lwd = 3)
grid(NULL, NULL, col = "darkgray")


plot(mean_PCR_load, type = "b",
     main = "Mean PCR Load per Time Point",
     xlab = "Time Point",
     ylab = "Mean # Leishmania parasites / mL blood")


boxplot(leish$PCR_Load_TP1,
        leish$PCR_Load_TP2,
        leish$PCR_Load_TP3,
        leish$PCR_Load_TP4,
        leish$PCR_Load_TP5,
        leish$PCR_Load_TP6,
        leish$PCR_Load_TP7,
        ylim = c(0,1000),
        at = 1:7,
        main = "PCR Load per Time Point",
        xlab = "Time Point",
        ylab = "# Leishmania parasites / mL blood",
        names = 1:7)


par(mfrow=c(2,2))
hist(leish$PCR_Load_TP1, main = "TP1", 
     xlab = "# Leishmania parasites / mL blood ", 
     col = "gray")
hist(leish$PCR_Load_TP2, main = "TP2", 
     xlab = "# Leishmania parasites / mL blood ", 
     col = "gray")
hist(leish$PCR_Load_TP3, main = "TP3", 
     xlab = "# Leishmania parasites / mL blood ", 
     col = "gray")
hist(leish$PCR_Load_TP4, main = "TP4", 
     xlab = "# Leishmania parasites / mL blood ", 
     col = "gray")
hist(leish$PCR_Load_TP5, main = "TP5", 
     xlab = "# Leishmania parasites / mL blood ", 
     col = "gray")
hist(leish$PCR_Load_TP6, main = "TP6", 
     xlab = "# Leishmania parasites / mL blood ", 
     col = "gray")
hist(leish$PCR_Load_TP7, main = "TP7", 
     xlab = "# Leishmania parasites / mL blood ", 
     col = "gray")
```

\ 

# MCMC Analysis and Results

\ 

## Scaled Datasets (Pathogen Load, Antibodies, Disease Status, and Covariates)

Since the data set contains large values for pathogen load over time, then original values were scaled by using a modified version of the original scale. We used #Parasites (in fifty thousands) / mL of blood. In the case of the SLA ELISA OD Ratio (antibody levels), we scaled by 10. The reason for scaling was only to solve some computational issues. 

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#### DATA ===========================================================
dt_P <- as.matrix(leish[,c("PCR_Load_TP1","PCR_Load_TP2", 
                           "PCR_Load_TP3", "PCR_Load_TP4",
                           "PCR_Load_TP5", "PCR_Load_TP6",
                           "PCR_Load_TP7")])
dt_D <- as.matrix(leish[,c("DStatus_TP1","DStatus_TP2", 
                           "DStatus_TP3", "DStatus_TP4",
                           "DStatus_TP5", "DStatus_TP6",
                           "DStatus_TP7")])
dt_A <- as.matrix(leish[,c("SLA.ELISA.OD.ratio_TP1",
                           "SLA.ELISA.OD.ratio_TP2", 
                           "SLA.ELISA.OD.ratio_TP3", 
                           "SLA.ELISA.OD.ratio_TP4",
                           "SLA.ELISA.OD.ratio_TP5",
                           "SLA.ELISA.OD.ratio_TP6",
                           "SLA.ELISA.OD.ratio_TP7")])

N = nrow(leish)
Time = ncol(dt_P)

dt_age <- data.frame(matrix(data = rep(leish$Age_group,Time), 
                            ncol = ncol(dt_P), nrow = nrow(dt_P)))
dt_snap <- data.frame(matrix(data = rep(leish$Ever.SNAP.Positive,Time), 
                             ncol = ncol(dt_P), nrow = nrow(dt_P)))
dt_dpp <- data.frame(matrix(data = rep(leish$DPP_Enroll,Time), 
                            ncol = ncol(dt_P), nrow = nrow(dt_P)))
dt_trt <- data.frame(matrix(data = rep(leish$treatment_group,Time), 
                            ncol = ncol(dt_P), nrow = nrow(dt_P)))

dt_age1 <- data.frame(matrix(data = as.numeric(rep(leish$Age_group,Time)), 
                            ncol = ncol(dt_P), nrow = nrow(dt_P)))
dt_snap1 <- data.frame(matrix(data = as.numeric(rep(leish$Ever.SNAP.Positive,Time)), 
                             ncol = ncol(dt_P), nrow = nrow(dt_P)))
dt_dpp1 <- data.frame(matrix(data = rep(leish$DPP_Enroll,Time), 
                            ncol = ncol(dt_P), nrow = nrow(dt_P)))
dt_trt1 <- data.frame(matrix(data = as.numeric(rep(leish$treatment_group,Time)), 
                            ncol = ncol(dt_P), nrow = nrow(dt_P)))
#write.csv(dt_P, file = "dt_P.csv", col.names = TRUE, row.names = FALSE, na = "")
#write.csv(dt_D, file = "dt_D.csv", col.names = TRUE, row.names = FALSE, na = "")
#write.csv(dt_A, file = "dt_A.csv", col.names = TRUE, row.names = FALSE, na = "")
write.csv(dt_age1, file = "dt_age.csv", col.names = TRUE, row.names = FALSE, na = "")
write.csv(dt_snap1, file = "dt_snap.csv", col.names = TRUE, row.names = FALSE, na = "")
write.csv(dt_dpp1, file = "dt_dpp.csv", col.names = TRUE, row.names = FALSE, na = "")
write.csv(dt_trt1, file = "dt_trt.csv", col.names = TRUE, row.names = FALSE, na = "")


# Scale Pathogen load (divide by 50,000; which is 5e+04) 
# Antibodies (divide by 10)
# DPP, in Covariates (divide by 200)
generalmaxP = 5e+04
generalmaxA = 10
generalmaxDPP = 200

dt_P = dt_P/generalmaxP
dt_A = dt_A/generalmaxA
dt_X <- cbind(1, leish$Age_group, 
              leish$Ever.SNAP.Positive,
              leish$DPP_Enroll/generalmaxDPP,
              leish$treatment_group)

otherdt_X <- cbind(1, 
              as.numeric(leish$Age_group)-1, 
              as.numeric(leish$Ever.SNAP.Positive)-1,
              leish$DPP_Enroll/generalmaxDPP,
              as.numeric(leish$treatment_group)-1)


#write.csv(dt_X, file = "dt_trt.csv", col.names = TRUE, row.names = FALSE, na = "")
#write.csv(otherdt_X, file = "dt_X.csv", col.names = TRUE, row.names = FALSE, na = "")

# function for multinomial
val_multinom <- function(y){
  dt <- data.frame(matrix(NA,ncol = 4))
  for(i in 1:length(y)){
    dt[i,] <- if(y[i] == 1){c(1,0,0,0)} 
    else if(y[i] == 2){c(0,1,0,0)} 
    else if(y[i] == 3){c(0,0,1,0)} 
    else if(y[i] == 4){c(0,0,0,1)} 
    else {NA} # in case D status is missing
  }
  dt
}

# entries in matrix that needs to be ignored due to removing of subjects
# dogs removed: 758, 761, 743
total_entries = N*Time
entries = dt_D
entries[1:total_entries] = 1:total_entries
ignore_entries = c(318,333,236,286,336)
ignore_whichID = c(18,33,36) 

# missing
missD = which(is.na(dt_D))[!(which(is.na(dt_D)) %in% ignore_entries)]
missA = which(is.na(dt_A))[!(which(is.na(dt_A)) %in% ignore_entries)]
missP = which(is.na(dt_P))[!(which(is.na(dt_P)) %in% ignore_entries)]

```

\ 

Pathogen Load (Scaled, divided by 50,000):
```{r, echo=FALSE, message=FALSE, warning=FALSE}
as.data.frame(dt_P)
```

\ 

Antibodies (Scaled, divided by 100):
```{r, echo=FALSE, message=FALSE, warning=FALSE}
as.data.frame(dt_A)
```

\ 

Disease Status:
```{r, echo=FALSE, message=FALSE, warning=FALSE}
as.data.frame(dt_D)
```

\ 

Intercept and Covariates (Age Group, Ever Snap Positive, DPP, Treatment Group):
```{r, echo=FALSE, message=FALSE, warning=FALSE}
temp = as.data.frame(dt_X)
colnames(temp) = c("Intercept", "Age_Group", "Ever_SNAP_Positive", "DPP_Results", "Treatment_Group")
temp
```

\ 

## Running MH-within Gibbs 

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#### Log(Full Conditionals) of Parameters 

betaP_FC <- function(x, sigma2p, alphap, D, P, A, X = dt_X){
  
  D1 <- apply(D,2,function(x){ifelse(x == 1, 1, 0)})
  D2 <- apply(D,2,function(x){ifelse(x == 2, 1, 0)})
  D3 <- apply(D,2,function(x){ifelse(x == 3, 1, 0)})
  
  sumterm = sum(dmvnorm(x, mean = rep(0,length(x)), 
                        sigma = diag(rep(1,length(x))), log = TRUE)) +
    sum(dnorm(P[,1], log = TRUE)) +
    sum(dnorm(A[,1], log = TRUE)) +
    sum(apply(val_multinom(D[,1]),1, function(y){dmultinom(y, 
                                                           prob = c(0.5,0.3,
                                                                    0.19,0.01), 
                                                           log = TRUE)}))
  for(i in 1:N){
    for(j in 2:Time){
      if((i == 18)&(j == 7)) next
      if((i == 33)&(j == 7)) next
      if((i == 36)&(j %in% c(5,6,7))) next
      M <- cbind(P[i,j-1], 
                 D1[i,j-1]*P[i,j-1], 
                 D2[i,j-1]*P[i,j-1], 
                 D3[i,j-1]*P[i,j-1],
                 if(j==2){0} else{D1[i,j-1]*(P[i,j-1]-P[i,j-2])},
                 if(j==2){0} else{D2[i,j-1]*(P[i,j-1]-P[i,j-2])},
                 if(j==2){0} else{D3[i,j-1]*(P[i,j-1]-P[i,j-2])},
                 if(j==2){0} else{D1[i,j-1]*(A[i,j-1]-A[i,j-2])}, 
                 if(j==2){0} else{D2[i,j-1]*(A[i,j-1]-A[i,j-2])}, 
                 if(j==2){0} else{D3[i,j-1]*(A[i,j-1]-A[i,j-2])})
      etap <- (M %*% x) + (X[i,] %*% alphap)
      sumterm <- sumterm + dnorm(P[i,j], etap, sigma2p, log = TRUE)
    }
  }

  sumterm
}

sigma2p_FC <- function(x, betaP, alphap, D, P, A, X = dt_X){
  
  D1 <- apply(D,2,function(x){ifelse(x == 1, 1, 0)})
  D2 <- apply(D,2,function(x){ifelse(x == 2, 1, 0)})
  D3 <- apply(D,2,function(x){ifelse(x == 3, 1, 0)})

  sumterm = sum(dgamma(x, 1, 1,log = TRUE)) +
    sum(dnorm(P[,1], log = TRUE)) +
    sum(dnorm(A[,1], log = TRUE)) +
    sum(apply(val_multinom(D[,1]),1, function(y){dmultinom(y, 
                                                           prob = c(0.5,0.3,
                                                                    0.19,0.01), 
                                                           log = TRUE)}))
  
  for(i in 1:N){
    for(j in 2:Time){
      if((i == 18)&(j == 7)) next
      if((i == 33)&(j == 7)) next
      if((i == 36)&(j %in% c(5,6,7))) next
      M <- cbind(P[i,j-1], 
                 D1[i,j-1]*P[i,j-1], 
                 D2[i,j-1]*P[i,j-1], 
                 D3[i,j-1]*P[i,j-1],
                 if(j==2){0} else{D1[i,j-1]*(P[i,j-1]-P[i,j-2])},
                 if(j==2){0} else{D2[i,j-1]*(P[i,j-1]-P[i,j-2])},
                 if(j==2){0} else{D3[i,j-1]*(P[i,j-1]-P[i,j-2])},
                 if(j==2){0} else{D1[i,j-1]*(A[i,j-1]-A[i,j-2])}, 
                 if(j==2){0} else{D2[i,j-1]*(A[i,j-1]-A[i,j-2])}, 
                 if(j==2){0} else{D3[i,j-1]*(A[i,j-1]-A[i,j-2])})
      etap <- (M %*% betaP) + (X[i,] %*% alphap)
      sumterm <- sumterm + dnorm(P[i,j], etap, x,log = TRUE)
    }
  }

  sumterm
}


alphap_FC <- function(x, betaP, sigma2p, D, P, A, X = dt_X){
  
  D1 <- apply(D,2,function(x){ifelse(x == 1, 1, 0)})
  D2 <- apply(D,2,function(x){ifelse(x == 2, 1, 0)})
  D3 <- apply(D,2,function(x){ifelse(x == 3, 1, 0)})

  sumterm = sum(dmvnorm(x, mean = rep(0,ncol(X)), 
                        sigma = diag(rep(1,ncol(X))), log = TRUE)) +
    sum(dnorm(P[,1], log = TRUE)) +
    sum(dnorm(A[,1], log = TRUE)) +
    sum(apply(val_multinom(D[,1]),1, function(y){dmultinom(y, 
                                                           prob = c(0.5,0.3,
                                                                    0.19,0.01), 
                                                           log = TRUE)}))
  for(i in 1:N){
    for(j in 2:Time){
      if((i == 18)&(j == 7)) next
      if((i == 33)&(j == 7)) next
      if((i == 36)&(j %in% c(5,6,7))) next
      M <- cbind(P[i,j-1], 
                 D1[i,j-1]*P[i,j-1], 
                 D2[i,j-1]*P[i,j-1], 
                 D3[i,j-1]*P[i,j-1],
                 if(j==2){0} else{D1[i,j-1]*(P[i,j-1]-P[i,j-2])},
                 if(j==2){0} else{D2[i,j-1]*(P[i,j-1]-P[i,j-2])},
                 if(j==2){0} else{D3[i,j-1]*(P[i,j-1]-P[i,j-2])},
                 if(j==2){0} else{D1[i,j-1]*(A[i,j-1]-A[i,j-2])}, 
                 if(j==2){0} else{D2[i,j-1]*(A[i,j-1]-A[i,j-2])}, 
                 if(j==2){0} else{D3[i,j-1]*(A[i,j-1]-A[i,j-2])})
      etap <- (M %*% betaP) + (X[i,] %*% x)
      sumterm <- sumterm + dnorm(P[i,j], etap, sigma2p,log = TRUE)
    }
  }

  sumterm
}


betaA_FC <- function(x, sigma2A, alphaA, D, P, A, X = dt_X){
  
  D1 <- apply(D,2,function(x){ifelse(x == 1, 1, 0)})
  D2 <- apply(D,2,function(x){ifelse(x == 2, 1, 0)})
  D3 <- apply(D,2,function(x){ifelse(x == 3, 1, 0)})

  sumterm = sum(dmvnorm(x, mean = rep(0,length(x)), 
                        sigma = diag(rep(1,length(x))), 
                        log = TRUE)) +
    sum(dnorm(P[,1], log = TRUE)) +
    sum(dnorm(A[,1], log = TRUE)) +
    sum(apply(val_multinom(D[,1]),1, function(y){dmultinom(y, 
                                                           prob = c(0.5,0.3,
                                                                    0.19,0.01), 
                                                           log = TRUE)}))
  for(i in 1:N){
    for(j in 2:Time){
      if((i == 18)&(j == 7)) next
      if((i == 33)&(j == 7)) next
      if((i == 36)&(j %in% c(5,6,7))) next
      M <- cbind(P[i,j-1], 
                 A[i,j-1],
                 if(j==2){0} else{D1[i,j-1]*(P[i,j-1]-P[i,j-2])},
                 if(j==2){0} else{D2[i,j-1]*(P[i,j-1]-P[i,j-2])},
                 if(j==2){0} else{D3[i,j-1]*(P[i,j-1]-P[i,j-2])},
                 if(j==2){0} else{D1[i,j-1]*(A[i,j-1]-A[i,j-2])}, 
                 if(j==2){0} else{D2[i,j-1]*(A[i,j-1]-A[i,j-2])}, 
                 if(j==2){0} else{D3[i,j-1]*(A[i,j-1]-A[i,j-2])})
      etaA <- (M %*% x) + (X[i,] %*% alphaA)
      sumterm <- sumterm + dnorm(A[i,j], etaA, sigma2A,log = TRUE)
    }
  }

  sumterm
}


alphaA_FC <- function(x, sigma2A, betaA, D, P, A, X = dt_X){
  
  D1 <- apply(D,2,function(x){ifelse(x == 1, 1, 0)})
  D2 <- apply(D,2,function(x){ifelse(x == 2, 1, 0)})
  D3 <- apply(D,2,function(x){ifelse(x == 3, 1, 0)})

  sumterm = sum(dmvnorm(x, mean = rep(0,ncol(X)), 
                        sigma = diag(rep(1,ncol(X))), 
                        log = TRUE)) +
    sum(dnorm(P[,1], log = TRUE)) +
    sum(dnorm(A[,1], log = TRUE)) +
    sum(apply(val_multinom(D[,1]),1, function(y){dmultinom(y, 
                                                           prob = c(0.5,0.3,
                                                                    0.19,0.01), 
                                                           log = TRUE)}))
  for(i in 1:N){
    for(j in 2:Time){
      if((i == 18)&(j == 7)) next
      if((i == 33)&(j == 7)) next
      if((i == 36)&(j %in% c(5,6,7))) next
      M <- cbind(P[i,j-1], 
                 A[i,j-1],
                 if(j==2){0} else{D1[i,j-1]*(P[i,j-1]-P[i,j-2])},
                 if(j==2){0} else{D2[i,j-1]*(P[i,j-1]-P[i,j-2])},
                 if(j==2){0} else{D3[i,j-1]*(P[i,j-1]-P[i,j-2])},
                 if(j==2){0} else{D1[i,j-1]*(A[i,j-1]-A[i,j-2])}, 
                 if(j==2){0} else{D2[i,j-1]*(A[i,j-1]-A[i,j-2])}, 
                 if(j==2){0} else{D3[i,j-1]*(A[i,j-1]-A[i,j-2])})
      etaA <- (M %*% betaA) + (X[i,] %*% x)
      sumterm <- sumterm + dnorm(A[i,j], etaA, sigma2A,log = TRUE)
    }
  }

  sumterm
}


sigma2A_FC <- function(x, betaA, alphaA, D, P, A, X = dt_X){
  
  D1 <- apply(D,2,function(x){ifelse(x == 1, 1, 0)})
  D2 <- apply(D,2,function(x){ifelse(x == 2, 1, 0)})
  D3 <- apply(D,2,function(x){ifelse(x == 3, 1, 0)})

  sumterm = sum(dgamma(x, 1, 1,log = TRUE)) +
    sum(dnorm(P[,1], log = TRUE)) +
    sum(dnorm(A[,1], log = TRUE)) +
    sum(apply(val_multinom(D[,1]),1, function(y){dmultinom(y, 
                                                           prob = c(0.5,0.3,
                                                                    0.19,0.01), 
                                                           log = TRUE)}))
  for(i in 1:N){
    for(j in 2:Time){
      if((i == 18)&(j == 7)) next
      if((i == 33)&(j == 7)) next
      if((i == 36)&(j %in% c(5,6,7))) next
      M <- cbind(P[i,j-1], 
                 A[i,j-1],
                 if(j==2){0} else{D1[i,j-1]*(P[i,j-1]-P[i,j-2])},
                 if(j==2){0} else{D2[i,j-1]*(P[i,j-1]-P[i,j-2])},
                 if(j==2){0} else{D3[i,j-1]*(P[i,j-1]-P[i,j-2])},
                 if(j==2){0} else{D1[i,j-1]*(A[i,j-1]-A[i,j-2])}, 
                 if(j==2){0} else{D2[i,j-1]*(A[i,j-1]-A[i,j-2])}, 
                 if(j==2){0} else{D3[i,j-1]*(A[i,j-1]-A[i,j-2])})
      etaA <- (M %*% betaA) + (X[i,] %*% alphaA)
      sumterm <- sumterm + dnorm(A[i,j], etaA, x, log = TRUE)
    }
  }

  sumterm
}


betaD1_FC <- function(x, betaD2, betaD3, alphaD, D, P, A, X = dt_X){
  
  D1 <- apply(D,2,function(x){ifelse(x == 1, 1, 0)})
  D2 <- apply(D,2,function(x){ifelse(x == 2, 1, 0)})
  D3 <- apply(D,2,function(x){ifelse(x == 3, 1, 0)})

  sumterm = sum(dmvnorm(x, mean = rep(0,length(x)), 
                        sigma = diag(rep(1,length(x))), log = TRUE)) +
    sum(dnorm(P[,1], log = TRUE)) +
    sum(dnorm(A[,1], log = TRUE)) +
    sum(apply(val_multinom(D[,1]),1, function(y){dmultinom(y, 
                                                           prob = c(0.5,0.3,
                                                                    0.19,0.01), 
                                                           log = TRUE)}))
  for(i in 1:N){
    for(j in 2:Time){
      
      if((i == 18)&(j == 7)) next
      if((i == 33)&(j == 7)) next
      if((i == 36)&(j %in% c(5,6,7))) next
      
      M1 <- cbind(D1[i,j-1], D1[i,j-1]*P[i,j-1], D1[i,j-1]*A[i,j-1])
      M2 <- cbind(D2[i,j-1], D2[i,j-1]*P[i,j-1], D2[i,j-1]*A[i,j-1])
      M3 <- cbind(D3[i,j-1], D3[i,j-1]*P[i,j-1], D3[i,j-1]*A[i,j-1])
      
      num1 <- exp((M1 %*% x) + (X[i,] %*% alphaD))
      den1 <- 1 + exp((M1 %*% x) + (X[i,] %*% alphaD)) + 
        exp((M2 %*% betaD2) + (X[i,] %*% alphaD)) + 
        exp((M3 %*% betaD3) + (X[i,] %*% alphaD))
      piD1 <- num1/den1
      
      num2 <- exp((M2 %*% betaD2) + (X[i,] %*% alphaD))
      den2 <- 1 + exp((M1 %*% x) + (X[i,] %*% alphaD)) + 
        exp((M2 %*% betaD2) + (X[i,] %*% alphaD)) + 
        exp((M3 %*% betaD3) + (X[i,] %*% alphaD))
      piD2 <- num2/den2
      
      num3 <- exp((M3 %*% betaD3) + (X[i,] %*% alphaD))
      den3 <- 1 + exp((M1 %*% x) + (X[i,] %*% alphaD)) + 
        exp((M2 %*% betaD2) + (X[i,] %*% alphaD)) + 
        exp((M3 %*% betaD3) + (X[i,] %*% alphaD))
      piD3 <- num3/den3
      
      piD4 <- 1 - piD1 - piD2 - piD3
      
      sumterm <- sumterm + dmultinom(val_multinom(D[i,j]), 
                                     prob = c(piD1,piD2,piD3,piD4), log = TRUE)
    }
  }

  sumterm
}


betaD2_FC <- function(x, betaD1, betaD3, alphaD, D, P, A, X = dt_X){
  
  D1 <- apply(D,2,function(x){ifelse(x == 1, 1, 0)})
  D2 <- apply(D,2,function(x){ifelse(x == 2, 1, 0)})
  D3 <- apply(D,2,function(x){ifelse(x == 3, 1, 0)})

  sumterm = sum(dmvnorm(x, mean = rep(0,length(x)), 
                        sigma = diag(rep(1,length(x))), log = TRUE)) +
    sum(dnorm(P[,1], log = TRUE)) +
    sum(dnorm(A[,1], log = TRUE)) +
    sum(apply(val_multinom(D[,1]),1, function(y){dmultinom(y, 
                                                           prob = c(0.5,0.3,
                                                                    0.19,0.01), 
                                                           log = TRUE)}))
  for(i in 1:N){
    for(j in 2:Time){
      
      if((i == 18)&(j == 7)) next
      if((i == 33)&(j == 7)) next
      if((i == 36)&(j %in% c(5,6,7))) next
      
      M1 <- cbind(D1[i,j-1], D1[i,j-1]*P[i,j-1], D1[i,j-1]*A[i,j-1])
      M2 <- cbind(D2[i,j-1], D2[i,j-1]*P[i,j-1], D2[i,j-1]*A[i,j-1])
      M3 <- cbind(D3[i,j-1], D3[i,j-1]*P[i,j-1], D3[i,j-1]*A[i,j-1])
      
      num1 <- exp((M1 %*% betaD1) + (X[i,] %*% alphaD))
      den1 <- 1 + exp((M1 %*% betaD1) + (X[i,] %*% alphaD)) + 
        exp((M2 %*% x) + (X[i,] %*% alphaD)) + 
        exp((M3 %*% betaD3) + (X[i,] %*% alphaD))
      piD1 <- num1/den1
      
      num2 <- exp((M2 %*% x) + (X[i,] %*% alphaD))
      den2 <- 1 + exp((M1 %*% betaD1) + (X[i,] %*% alphaD)) + 
        exp((M2 %*% x) + (X[i,] %*% alphaD)) + 
        exp((M3 %*% betaD3) + (X[i,] %*% alphaD))
      piD2 <- num2/den2
      
      num3 <- exp((M3 %*% betaD3) + (X[i,] %*% alphaD))
      den3 <- 1 + exp((M1 %*% betaD1) + (X[i,] %*% alphaD)) + 
        exp((M2 %*% x) + (X[i,] %*% alphaD)) + 
        exp((M3 %*% betaD3) + (X[i,] %*% alphaD))
      piD3 <- num3/den3
      
      piD4 <- 1 - piD1 - piD2 - piD3
      
      sumterm <- sumterm + dmultinom(val_multinom(D[i,j]), 
                                     prob = c(piD1,piD2,piD3,piD4), log = TRUE)
    }
  }

  sumterm
}


betaD3_FC <- function(x, betaD1, betaD2, alphaD, D, P, A, X = dt_X){
  
  D1 <- apply(D,2,function(x){ifelse(x == 1, 1, 0)})
  D2 <- apply(D,2,function(x){ifelse(x == 2, 1, 0)})
  D3 <- apply(D,2,function(x){ifelse(x == 3, 1, 0)})

  sumterm = sum(dmvnorm(x, mean = rep(0,length(x)), 
                        sigma = diag(rep(1,length(x))), log = TRUE)) +
    sum(dnorm(P[,1], log = TRUE)) +
    sum(dnorm(A[,1], log = TRUE)) +
    sum(apply(val_multinom(D[,1]),1, function(y){dmultinom(y, 
                                                           prob = c(0.5,0.3,
                                                                    0.19,0.01), 
                                                           log = TRUE)}))
  for(i in 1:N){
    for(j in 2:Time){
      
      if((i == 18)&(j == 7)) next
      if((i == 33)&(j == 7)) next
      if((i == 36)&(j %in% c(5,6,7))) next
      
      M1 <- cbind(D1[i,j-1], D1[i,j-1]*P[i,j-1], D1[i,j-1]*A[i,j-1])
      M2 <- cbind(D2[i,j-1], D2[i,j-1]*P[i,j-1], D2[i,j-1]*A[i,j-1])
      M3 <- cbind(D3[i,j-1], D3[i,j-1]*P[i,j-1], D3[i,j-1]*A[i,j-1])
      
      num1 <- exp((M1 %*% betaD1) + (X[i,] %*% alphaD))
      den1 <- 1 + exp((M1 %*% betaD1) + (X[i,] %*% alphaD)) + 
        exp((M2 %*% betaD2) + (X[i,] %*% alphaD)) + 
        exp((M3 %*% x) + (X[i,] %*% alphaD))
      piD1 <- num1/den1
      
      num2 <- exp((M2 %*% betaD2) + (X[i,] %*% alphaD))
      den2 <- 1 + exp((M1 %*% betaD1) + (X[i,] %*% alphaD)) + 
        exp((M2 %*% betaD2) + (X[i,] %*% alphaD)) + 
        exp((M3 %*% x) + (X[i,] %*% alphaD))
      piD2 <- num2/den2
      
      num3 <- exp((M3 %*% x) + (X[i,] %*% alphaD))
      den3 <- 1 + exp((M1 %*% betaD1) + (X[i,] %*% alphaD)) + 
        exp((M2 %*% betaD2) + (X[i,] %*% alphaD)) + 
        exp((M3 %*% x) + (X[i,] %*% alphaD))
      piD3 <- num3/den3
      
      piD4 <- 1 - piD1 - piD2 - piD3
      
      sumterm <- sumterm + dmultinom(val_multinom(D[i,j]), 
                                     prob = c(piD1,piD2,piD3,piD4), log = TRUE)
    }
  }

  sumterm
}


alphaD_FC <- function(x, betaD1, betaD2, betaD3, D, P, A, X = dt_X){
  
  D1 <- apply(D,2,function(x){ifelse(x == 1, 1, 0)})
  D2 <- apply(D,2,function(x){ifelse(x == 2, 1, 0)})
  D3 <- apply(D,2,function(x){ifelse(x == 3, 1, 0)})

  sumterm = sum(dmvnorm(x, mean = rep(0,ncol(X)), 
                        sigma = diag(rep(1,ncol(X))), log = TRUE)) +
    sum(dnorm(P[,1], log = TRUE)) +
    sum(dnorm(A[,1], log = TRUE)) +
    sum(apply(val_multinom(D[,1]),1, function(y){dmultinom(y, 
                                                           prob = c(0.5,0.3,
                                                                    0.19,0.01), 
                                                           log = TRUE)}))
  for(i in 1:N){
    for(j in 2:Time){
      
      if((i == 18)&(j == 7)) next
      if((i == 33)&(j == 7)) next
      if((i == 36)&(j %in% c(5,6,7))) next
      
      M1 <- cbind(D1[i,j-1], D1[i,j-1]*P[i,j-1], D1[i,j-1]*A[i,j-1])
      M2 <- cbind(D2[i,j-1], D2[i,j-1]*P[i,j-1], D2[i,j-1]*A[i,j-1])
      M3 <- cbind(D3[i,j-1], D3[i,j-1]*P[i,j-1], D3[i,j-1]*A[i,j-1])
      
      num1 <- exp((M1 %*% betaD1) + (X[i,] %*% x))
      den1 <- 1 + exp((M1 %*% betaD1) + (X[i,] %*% x)) + 
        exp((M2 %*% betaD2) + (X[i,] %*% x)) + 
        exp((M3 %*% betaD3) + (X[i,] %*% x))
      piD1 <- num1/den1
      
      num2 <- exp((M2 %*% betaD2) + (X[i,] %*% x))
      den2 <- 1 + exp((M1 %*% betaD1) + (X[i,] %*% x)) + 
        exp((M2 %*% betaD2) + (X[i,] %*% x)) + 
        exp((M3 %*% betaD3) + (X[i,] %*% x))
      piD2 <- num2/den2
      
      num3 <- exp((M3 %*% betaD3) + (X[i,] %*% x))
      den3 <- 1 + exp((M1 %*% betaD1) + (X[i,] %*% x)) + 
        exp((M2 %*% betaD2) + (X[i,] %*% x)) + 
        exp((M3 %*% betaD3) + (X[i,] %*% x))
      piD3 <- num3/den3
      
      piD4 <- 1 - piD1 - piD2 - piD3
      
      sumterm <- sumterm + dmultinom(val_multinom(D[i,j]), 
                                     prob = c(piD1,piD2,piD3,piD4), log = TRUE)
      
    }
  }

  sumterm
}


D_FC <- function(D, betaP, alphap, sigma2p, betaD1, betaD2, 
                 betaD3, alphaD, P, A, X = dt_X){
  
  D1 <- apply(D,2,function(x){ifelse(x == 1, 1, 0)})
  D2 <- apply(D,2,function(x){ifelse(x == 2, 1, 0)})
  D3 <- apply(D,2,function(x){ifelse(x == 3, 1, 0)})
  
  sumterm = sum(dnorm(P[,1], log = TRUE)) +
    sum(dnorm(A[,1], log = TRUE)) +
    sum(apply(val_multinom(D[,1]),1, function(y){dmultinom(y, 
                                                           prob = c(0.5,0.3,
                                                                    0.19,0.01), 
                                                           log = TRUE)}))
  
  for(i in 1:N){
    for(j in 2:Time){
      
      if((i == 18)&(j == 7)) next
      if((i == 33)&(j == 7)) next
      if((i == 36)&(j %in% c(5,6,7))) next
      
      M <- cbind(P[i,j-1], 
                 D1[i,j-1]*P[i,j-1], 
                 D2[i,j-1]*P[i,j-1], 
                 D3[i,j-1]*P[i,j-1],
                 if(j==2){0} else{D1[i,j-1]*(P[i,j-1]-P[i,j-2])},
                 if(j==2){0} else{D2[i,j-1]*(P[i,j-1]-P[i,j-2])},
                 if(j==2){0} else{D3[i,j-1]*(P[i,j-1]-P[i,j-2])},
                 if(j==2){0} else{D1[i,j-1]*(A[i,j-1]-A[i,j-2])}, 
                 if(j==2){0} else{D2[i,j-1]*(A[i,j-1]-A[i,j-2])}, 
                 if(j==2){0} else{D3[i,j-1]*(A[i,j-1]-A[i,j-2])})
      etap <- (M %*% betaP) + (X[i,] %*% alphap)
      
      M1 <- cbind(D1[i,j-1], D1[i,j-1]*P[i,j-1], D1[i,j-1]*A[i,j-1])
      M2 <- cbind(D2[i,j-1], D2[i,j-1]*P[i,j-1], D2[i,j-1]*A[i,j-1])
      M3 <- cbind(D3[i,j-1], D3[i,j-1]*P[i,j-1], D3[i,j-1]*A[i,j-1])
      
      num1 <- exp((M1 %*% betaD1) + (X[i,] %*% alphaD))
      den1 <- 1 + exp((M1 %*% betaD1) + (X[i,] %*% alphaD)) + 
        exp((M2 %*% betaD2) + (X[i,] %*% alphaD)) + 
        exp((M3 %*% betaD3) + (X[i,] %*% alphaD))
      piD1 <- num1/den1
      
      num2 <- exp((M2 %*% betaD2) + (X[i,] %*% alphaD))
      den2 <- 1 + exp((M1 %*% betaD1) + (X[i,] %*% alphaD)) + 
        exp((M2 %*% betaD2) + (X[i,] %*% alphaD)) + 
        exp((M3 %*% betaD3) + (X[i,] %*% alphaD))
      piD2 <- num2/den2
      
      num3 <- exp((M3 %*% betaD3) + (X[i,] %*% alphaD))
      den3 <- 1 + exp((M1 %*% betaD1) + (X[i,] %*% alphaD)) + 
        exp((M2 %*% betaD2) + (X[i,] %*% alphaD)) + 
        exp((M3 %*% betaD3) + (X[i,] %*% alphaD))
      piD3 <- num3/den3
      
      piD4 <- 1 - piD1 - piD2 - piD3
      
      sumterm <- sumterm + dnorm(P[i,j], etap, sigma2p, log = TRUE) + 
        dmultinom(val_multinom(D[i,j]), prob = c(piD1,piD2,piD3,piD4), log = TRUE)
    }
  }
  sumterm  
} 


P_FC <- function(P, betaP, sigma2p, alphap, D, A, X = dt_X){
  
  D1 <- apply(D,2,function(x){ifelse(x == 1, 1, 0)})
  D2 <- apply(D,2,function(x){ifelse(x == 2, 1, 0)})
  D3 <- apply(D,2,function(x){ifelse(x == 3, 1, 0)})
  
  sumterm = sum(dnorm(P[,1], log = TRUE)) +
    sum(dnorm(A[,1], log = TRUE)) +
    sum(apply(val_multinom(D[,1]),1, function(y){dmultinom(y, 
                                                           prob = c(0.5,0.3,
                                                                    0.19,0.01), 
                                                           log = TRUE)}))
  for(i in 1:N){
    for(j in 2:Time){
      
      if((i == 18)&(j == 7)) next
      if((i == 33)&(j == 7)) next
      if((i == 36)&(j %in% c(5,6,7))) next
      
      M <- cbind(P[i,j-1], 
                 D1[i,j-1]*P[i,j-1], 
                 D2[i,j-1]*P[i,j-1], 
                 D3[i,j-1]*P[i,j-1],
                 if(j==2){0} else{D1[i,j-1]*(P[i,j-1]-P[i,j-2])},
                 if(j==2){0} else{D2[i,j-1]*(P[i,j-1]-P[i,j-2])},
                 if(j==2){0} else{D3[i,j-1]*(P[i,j-1]-P[i,j-2])},
                 if(j==2){0} else{D1[i,j-1]*(A[i,j-1]-A[i,j-2])}, 
                 if(j==2){0} else{D2[i,j-1]*(A[i,j-1]-A[i,j-2])}, 
                 if(j==2){0} else{D3[i,j-1]*(A[i,j-1]-A[i,j-2])})
      etap <- (M %*% betaP) + (X[i,] %*% alphap)
      sumterm <- sumterm + dnorm(P[i,j], etap, sigma2p,log = TRUE)
    }
  }

  sumterm
}


A_FC <- function(A, betaA, alphaA, sigma2A, D, P, X = dt_X){
  
  D1 <- apply(D,2,function(x){ifelse(x == 1, 1, 0)})
  D2 <- apply(D,2,function(x){ifelse(x == 2, 1, 0)})
  D3 <- apply(D,2,function(x){ifelse(x == 3, 1, 0)})

  sumterm = sum(dnorm(P[,1], log = TRUE)) +
    sum(dnorm(A[,1], log = TRUE)) +
    sum(apply(val_multinom(D[,1]),1, function(y){dmultinom(y, 
                                                           prob = c(0.5,0.3,
                                                                    0.19,0.01), 
                                                           log = TRUE)}))
  for(i in 1:N){
    for(j in 2:Time){
      
      if((i == 18)&(j == 7)) next
      if((i == 33)&(j == 7)) next
      if((i == 36)&(j %in% c(5,6,7))) next
      
      M <- cbind(P[i,j-1], 
                 A[i,j-1],
                 if(j==2){0} else{D1[i,j-1]*(P[i,j-1]-P[i,j-2])},
                 if(j==2){0} else{D2[i,j-1]*(P[i,j-1]-P[i,j-2])},
                 if(j==2){0} else{D3[i,j-1]*(P[i,j-1]-P[i,j-2])},
                 if(j==2){0} else{D1[i,j-1]*(A[i,j-1]-A[i,j-2])}, 
                 if(j==2){0} else{D2[i,j-1]*(A[i,j-1]-A[i,j-2])}, 
                 if(j==2){0} else{D3[i,j-1]*(A[i,j-1]-A[i,j-2])})
      etaA <- (M %*% betaA) + (X[i,] %*% alphaA)
      sumterm <- sumterm + dnorm(A[i,j], etaA, sigma2A, log = TRUE)
    }
  }

  sumterm
}


#### Proposal Function, Single MH Updater =======================

r_function <- function(from, sdtuning = 1) {
  mvrnorm(1, mu=from, Sigma = diag(sdtuning, length(from)))
}

d_function <- function(from, to, sdtuning = 1) {
  dmvnorm(to, mean=from, sigma = diag(sdtuning, length(from)), log = TRUE)
}

dmvnorm_function <- function(from, to, sdtuning = 1) {
  to = as.vector(to)[-ignore_entries]
  from = as.vector(from)[-ignore_entries]
  dmvnorm(to, 
          mean = from, 
          sigma = diag(sdtuning, length(from)),
          log = TRUE)
}

propose_D <- function(D, missing_ind_D, propose_proportion = 0.1){
  D_proposed <- D
  to_propose <- sample(missing_ind_D, 
                       size = ceiling(propose_proportion*length(missing_ind_D)))
  proposal <- sample(1:4, size = length(to_propose), replace = TRUE)
  D_proposed[to_propose] <- proposal
  D_proposed
}

draw_D <- function(D, missing_ind_D, propose_proportion = 0.1, ...){
  
  D_proposed <- propose_D(D, missing_ind_D, propose_proportion)
  lpost_current <- D_FC(D, ...)
  lpost_proposed <-  D_FC(D_proposed, ...)
  lacc_prob <- lpost_proposed - lpost_current
  
  if (is.finite(lacc_prob) && log(runif(1,0,1)) < lacc_prob){
    return(D_proposed)
  }
  return(D)
}

newpropose_D <- function(D, missing_ind_D, propose_proportion = 0.1){
  D_proposed <- D
  to_propose <- sample(missing_ind_D, 
                       size = ceiling(propose_proportion*length(missing_ind_D)))
  proposal <- NULL
  for(i in 1:length(to_propose)){
    if(D_proposed[to_propose][i] == 1){
      proposal[i] = which.max(rmultinom(1,1,prob = c(1/2,1/2)))
    } else if(D_proposed[to_propose][i] == 2){
      proposal[i] = which.max(rmultinom(1,1,prob = c(1/3,1/3,1/3)))
    } else if(D_proposed[to_propose][i] == 3){
      proposal[i] = which.max(rmultinom(1,1,prob = c(0,1/3,1/3,1/3)))
    } else {proposal[i] = which.max(rmultinom(1,1,prob = c(0,0,1/2,1/2)))} 
  }
  D_proposed[to_propose] <- proposal
  D_proposed
}

newdraw_D <- function(D, missing_ind_D, propose_proportion = 0.1, ...){
  D_proposed <- newpropose_D(D, missing_ind_D, propose_proportion)
  lpost_current <- D_FC(D, ...)
  lpost_proposed <-  D_FC(D_proposed, ...)
  lacc_prob <- lpost_proposed - lpost_current
  
  if (is.finite(lacc_prob) && log(runif(1,0,1)) < lacc_prob){
    return(D_proposed)
  }
  return(D)
}


propose_A <- function(A, missing_ind_A, propose_proportion = 0.1, sdtuning = 0.1){
  A_proposed <- A
  to_propose <- sample(missing_ind_A, 
                       size = ceiling(propose_proportion*length(missing_ind_A)))
  proposal <- rnorm(rep(1, length(to_propose)), mean = 0, sd = sdtuning)
  A_proposed[to_propose] <- proposal
  A_proposed
}

draw_A <- function(A, missing_ind_A, propose_proportion = 0.1, sdtuning = 0.1, ...){
  
  A_proposed <- propose_A(A, missing_ind_A, propose_proportion, sdtuning)
  a <- sum(A_FC(A_proposed, ...) - A_FC(A, ...)) + 
    dmvnorm_function(A_proposed, A) -
    dmvnorm_function(A, A_proposed)
  
  if (is.finite(a) && log(runif(1,0,1)) < a){
    return(A_proposed)
  }
  return(A)
}


propose_P <- function(P, missing_ind_P, propose_proportion = 0.1, sdtuning = 0.1){
  P_proposed <- P
  to_propose <- sample(missing_ind_P, 
                       size = ceiling(propose_proportion*length(missing_ind_P)))
  proposal <- rnorm(rep(1, length(to_propose)), mean = 0, sd = sdtuning)
  P_proposed[to_propose] <- proposal
  P_proposed
}

draw_P <- function(P, missing_ind_P, propose_proportion = 0.1, sdtuning = 0.1, ...){
  
  #browser()
  P_proposed <- propose_P(P, missing_ind_P, propose_proportion, sdtuning)
  a <- sum(P_FC(P_proposed, ...) - P_FC(P, ...)) + 
    dmvnorm_function(P_proposed, P) -
    dmvnorm_function(P, P_proposed)
  
  if (is.finite(a) && log(runif(1,0,1)) < a){
    return(P_proposed)
  }
  return(P)
}


MH_Update <- function(x0, logFC, dfun, rfun, sdtuning = 1, ...){
  
  x1 <- rfun(x0, sdtuning = sdtuning)
  a <- suppressWarnings(sum((logFC(x=x1,...) - logFC(x=x0,...)) + 
                              dfun(x1, x0) - dfun(x0, x1)))
  u <- log(runif(1,0,1))
  if (is.finite(a) && (!is.nan(a)) && (u < a)){
    return(x1)
    }
  return(x0)
  
}


draw_onebeta <- function(x0, logFC, dfun, rfun, sdtuning = 1, idx = 1, ...){
  
  temp <- rfun(x0[idx], sdtuning = sdtuning)
  x1 <- x0
  x1[idx] <- temp

  a <- suppressWarnings(sum((logFC(x=x1,...) - logFC(x=x0,...)) + 
                              dfun(x1, x0) - dfun(x0, x1)))
  
  u <- log(runif(1,0,1))
  if (is.finite(a) && (!is.nan(a)) && (u < a)){
    return(x1)
    }
  return(x0)
  
}

  
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
####  MH-Gibbs =======================

MH_Gibbs <- function(n_iter, dataD = dt_D, dataA = dt_A, 
                     dataP = dt_P, filename = "chain.csv") {
  
  # missing indicators
  missing_ind_D <- which(is.na(dataD))[!(which(is.na(dataD)) %in% ignore_entries)]
  missing_ind_P <- which(is.na(dataP))[!(which(is.na(dataP)) %in% ignore_entries)]
  missing_ind_A <- which(is.na(dataA))[!(which(is.na(dataA)) %in% ignore_entries)]
  
  # pick initial values
  betaP_0 = rnorm(10,0,0.1)
  sigma2p_0 = rgamma(1,100,1000)
  alphap_0 = rnorm(ncol(dt_X),0,0.1)
  betaA_0 = rnorm(8,0,0.1)
  sigma2A_0 = rgamma(1,100,1000)
  alphaA_0 = rnorm(ncol(dt_X),0,0.1)
  betaD1_0 = rnorm(3,0,0.1)
  betaD2_0 = rnorm(3,0,0.1)
  betaD3_0 = rnorm(3,0,0.1)
  alphaD_0 = rnorm(ncol(dt_X),0,0.1)
  D0_for_missing = sample(1:4, 
                          size = length(missD), 
                          replace = TRUE)
  P0_for_missing = rep(0,length(missP))
  A0_for_missing = rep(0,length(missA))
  
  # creating vectors and matrices
  sigma2p_vec <- sigma2A_vec <- rep(0,n_iter)
  betaP_mat <- matrix(0, nrow=n_iter, ncol=length(betaP_0))
  betaA_mat <- matrix(0, nrow=n_iter, ncol=length(betaA_0))
  betaD1_mat <- matrix(0, nrow=n_iter, ncol=length(betaD1_0))
  betaD2_mat <- matrix(0, nrow=n_iter, ncol=length(betaD2_0))
  betaD3_mat <- matrix(0, nrow=n_iter, ncol=length(betaD3_0))
  alphap_mat <- alphaA_mat <- alphaD_mat <- matrix(0, 
                                                   nrow=n_iter, 
                                                   ncol=length(alphap_0))
  D_mat <- matrix(0, nrow=n_iter, ncol=length(missing_ind_D))
  P_mat <- matrix(0, nrow=n_iter, ncol=length(missing_ind_P))
  A_mat <- matrix(0, nrow=n_iter, ncol=length(missing_ind_A))
  
  # setting initial
  D_mat[1,] <- D0_for_missing
  P_mat[1,] <- P0_for_missing
  A_mat[1,] <- A0_for_missing
  betaP_mat[1,] <- betaP_0
  betaA_mat[1,] <- betaA_0
  betaD1_mat[1,] <- betaD1_0
  betaD2_mat[1,] <- betaD2_0
  betaD3_mat[1,] <- betaD3_0
  alphap_mat[1,] <- alphap_0
  alphaA_mat[1,] <- alphaA_0
  alphaD_mat[1,] <- alphaD_0
  sigma2p_vec[1] <- sigma2p_0
  sigma2A_vec[1] <- sigma2A_0
  
  ## adding updated (iterated) values to CSV file
  coef_names <- c(paste0("betaP_", 1:length(betaP_0)), "sigma2p",
                  paste0("betaA_", 1:length(betaA_0)), "sigma2A",
                  paste0("betaD1_", 1:length(betaD1_0)),
                  paste0("betaD2_", 1:length(betaD2_0)),
                  paste0("betaD3_", 1:length(betaD3_0)),
                  paste0("alphap_", 1:length(alphap_0)),
                  paste0("alphaA_", 1:length(alphaA_0)),
                  paste0("alphaD_", 1:length(alphaD_0)),
                  paste0(paste0("D_", (1:prod(dim(dt_D)))[missing_ind_D]), 
                         collapse = ","),
                  paste0(paste0("P_", (1:prod(dim(dt_P)))[missing_ind_P]), 
                         collapse = ","),
                  paste0(paste0("A_", (1:prod(dim(dt_A)))[missing_ind_A]), 
                         collapse = ","))
  cat(paste0(coef_names, collapse = ","), file = filename)
  cat("\n", file = filename, append=TRUE)
  init <- c(betaP_mat[1,], sigma2p_vec[1], betaA_mat[1,], sigma2A_vec[1], 
            betaD1_mat[1,], betaD2_mat[1,], betaD3_mat[1,], alphap_mat[1,], 
            alphaA_mat[1,], alphaD_mat[1,], D_mat[1,], P_mat[1,], A_mat[1,])
  cat(paste0(init, collapse = ","), file = filename, append=TRUE)
  cat("\n", file = filename, append=TRUE)
  
  D <- dataD; P <- dataP; A <- dataA
  D[missing_ind_D] <- D0_for_missing
  P[missing_ind_P] <- P0_for_missing
  A[missing_ind_A] <- A0_for_missing
  
  # loop for update parameters
  for (it in 2:n_iter) {
    
    bp1 <- draw_onebeta(x0 = betaP_mat[it-1,], 
                       logFC = betaP_FC, 
                       dfun = d_function, 
                       rfun = r_function,
                       sdtuning = 0.1, #previous: 0.05
                       sigma2p = sigma2p_vec[it-1], 
                       alphap = alphap_mat[it-1,],
                       idx = 1, 
                       D = D,
                       P = P,
                       A = A)
    betaP_mat[it,1] <- bp1[1]
    
    
    bp2 <- draw_onebeta(x0 = betaP_mat[it-1,], 
                       logFC = betaP_FC, 
                       dfun = d_function, 
                       rfun = r_function,
                       sdtuning = 4.5, #previous: 0.035
                       sigma2p = sigma2p_vec[it-1], 
                       alphap = alphap_mat[it-1,],
                       idx = 2, 
                       D = D,
                       P = P,
                       A = A)
    betaP_mat[it,2] <- bp2[2]
    
    
    bp3 <- draw_onebeta(x0 = betaP_mat[it-1,], 
                       logFC = betaP_FC, 
                       dfun = d_function, 
                       rfun = r_function,
                       sdtuning = 0.1, #previous: 0.035
                       sigma2p = sigma2p_vec[it-1], 
                       alphap = alphap_mat[it-1,],
                       idx = 3, 
                       D = D,
                       P = P,
                       A = A)
    betaP_mat[it,3] <- bp3[3]
    
    
    bp4 <- draw_onebeta(x0 = betaP_mat[it-1,], 
                       logFC = betaP_FC, 
                       dfun = d_function, 
                       rfun = r_function,
                       sdtuning = 4.95, #previous: 0.035
                       sigma2p = sigma2p_vec[it-1], 
                       alphap = alphap_mat[it-1,],
                       idx = 4, 
                       D = D,
                       P = P,
                       A = A)
    betaP_mat[it,4] <- bp4[4]
    
    
    
    bp5 <- draw_onebeta(x0 = betaP_mat[it-1,], 
                       logFC = betaP_FC, 
                       dfun = d_function, 
                       rfun = r_function,
                       sdtuning = 4.5, #previous: 0.035
                       sigma2p = sigma2p_vec[it-1], 
                       alphap = alphap_mat[it-1,],
                       idx = 5, 
                       D = D,
                       P = P,
                       A = A)
    betaP_mat[it,5] <- bp5[5]
    
    
    
    bp6 <- draw_onebeta(x0 = betaP_mat[it-1,], 
                       logFC = betaP_FC, 
                       dfun = d_function, 
                       rfun = r_function,
                       sdtuning = 0.05, #previous: 0.035
                       sigma2p = sigma2p_vec[it-1], 
                       alphap = alphap_mat[it-1,],
                       idx = 6, 
                       D = D,
                       P = P,
                       A = A)
    betaP_mat[it,6] <- bp6[6]
    
    
    
    bp7 <- draw_onebeta(x0 = betaP_mat[it-1,], 
                       logFC = betaP_FC, 
                       dfun = d_function, 
                       rfun = r_function,
                       sdtuning = 4.5, #previous: 0.035
                       sigma2p = sigma2p_vec[it-1], 
                       alphap = alphap_mat[it-1,],
                       idx = 7, 
                       D = D,
                       P = P,
                       A = A)
    betaP_mat[it,7] <- bp7[7]
    
    
    
    bp8 <- draw_onebeta(x0 = betaP_mat[it-1,], 
                       logFC = betaP_FC, 
                       dfun = d_function, 
                       rfun = r_function,
                       sdtuning = 0.01, #previous: 0.035
                       sigma2p = sigma2p_vec[it-1], 
                       alphap = alphap_mat[it-1,],
                       idx = 8, 
                       D = D,
                       P = P,
                       A = A)
    betaP_mat[it,8] <- bp8[8]
    
    
    
    bp9 <- draw_onebeta(x0 = betaP_mat[it-1,], 
                       logFC = betaP_FC, 
                       dfun = d_function, 
                       rfun = r_function,
                       sdtuning = 0.01, #previous: 0.035
                       sigma2p = sigma2p_vec[it-1], 
                       alphap = alphap_mat[it-1,],
                       idx = 9, 
                       D = D,
                       P = P,
                       A = A)
    betaP_mat[it,9] <- bp9[9]
    
    
    
    bp10 <- draw_onebeta(x0 = betaP_mat[it-1,], 
                       logFC = betaP_FC, 
                       dfun = d_function, 
                       rfun = r_function,
                       sdtuning = 0.15, #previous: 0.035
                       sigma2p = sigma2p_vec[it-1], 
                       alphap = alphap_mat[it-1,],
                       idx = 10, 
                       D = D,
                       P = P,
                       A = A)
    betaP_mat[it,10] <- bp10[10]
    
    
    
    sigma2p_result <- MH_Update(x0 = sigma2p_vec[it-1], 
                              logFC = sigma2p_FC, 
                              dfun = d_function, 
                              rfun = r_function,
                              sdtuning = 0.000001, #previous: 0.00001
                              betaP = betaP_mat[it-1,],
                              alphap = alphap_mat[it-1,],
                              D = D,
                              P = P,
                              A = A)
    sigma2p_vec[it] <- sigma2p_result
    
    
    ba1 <- draw_onebeta(x0 = betaA_mat[it-1,], 
                       logFC = betaA_FC, 
                       dfun = d_function, 
                       rfun = r_function,
                       sdtuning = 0.25, #previous: 0.035
                       sigma2A = sigma2A_vec[it-1], 
                       alphaA = alphaA_mat[it-1,],
                       idx = 1, 
                       D = D,
                       P = P,
                       A = A)
    betaA_mat[it,1] <- ba1[1]
    
    
    
    ba2 <- draw_onebeta(x0 = betaA_mat[it-1,], 
                       logFC = betaA_FC, 
                       dfun = d_function, 
                       rfun = r_function,
                       sdtuning = 0.005, #previous: 0.035
                       sigma2A = sigma2A_vec[it-1], 
                       alphaA = alphaA_mat[it-1,],
                       idx = 2, 
                       D = D,
                       P = P,
                       A = A)
    betaA_mat[it,2] <- ba2[2]
    
    
    
    ba3 <- draw_onebeta(x0 = betaA_mat[it-1,], 
                       logFC = betaA_FC, 
                       dfun = d_function, 
                       rfun = r_function,
                       sdtuning = 3, #previous: 0.035
                       sigma2A = sigma2A_vec[it-1], 
                       alphaA = alphaA_mat[it-1,],
                       idx = 3, 
                       D = D,
                       P = P,
                       A = A)
    betaA_mat[it,3] <- ba3[3]
    
    
    
    ba4 <- draw_onebeta(x0 = betaA_mat[it-1,], 
                       logFC = betaA_FC, 
                       dfun = d_function, 
                       rfun = r_function,
                       sdtuning = 0.25, #previous: 0.035
                       sigma2A = sigma2A_vec[it-1], 
                       alphaA = alphaA_mat[it-1,],
                       idx = 4, 
                       D = D,
                       P = P,
                       A = A)
    betaA_mat[it,4] <- ba4[4]
    
    
    
    ba5 <- draw_onebeta(x0 = betaA_mat[it-1,], 
                       logFC = betaA_FC, 
                       dfun = d_function, 
                       rfun = r_function,
                       sdtuning = 2.5, #previous: 0.035
                       sigma2A = sigma2A_vec[it-1], 
                       alphaA = alphaA_mat[it-1,],
                       idx = 5, 
                       D = D,
                       P = P,
                       A = A)
    betaA_mat[it,5] <- ba5[5]
    
    
    
    ba6 <- draw_onebeta(x0 = betaA_mat[it-1,], 
                       logFC = betaA_FC, 
                       dfun = d_function, 
                       rfun = r_function,
                       sdtuning = 0.05, #previous: 0.035
                       sigma2A = sigma2A_vec[it-1], 
                       alphaA = alphaA_mat[it-1,],
                       idx = 6, 
                       D = D,
                       P = P,
                       A = A)
    betaA_mat[it,6] <- ba6[6]
    
    
    
    ba7 <- draw_onebeta(x0 = betaA_mat[it-1,], 
                       logFC = betaA_FC, 
                       dfun = d_function, 
                       rfun = r_function,
                       sdtuning = 0.05, #previous: 0.035
                       sigma2A = sigma2A_vec[it-1], 
                       alphaA = alphaA_mat[it-1,],
                       idx = 7, 
                       D = D,
                       P = P,
                       A = A)
    betaA_mat[it,7] <- ba7[7]
    
    
    
    ba8 <- draw_onebeta(x0 = betaA_mat[it-1,], 
                       logFC = betaA_FC, 
                       dfun = d_function, 
                       rfun = r_function,
                       sdtuning = 0.65, #previous: 0.035
                       sigma2A = sigma2A_vec[it-1], 
                       alphaA = alphaA_mat[it-1,],
                       idx = 8, 
                       D = D,
                       P = P,
                       A = A)
    betaA_mat[it,8] <- ba8[8]
    
    
    
    sigma2A_result <- MH_Update(x0 = sigma2A_vec[it-1], 
                              logFC = sigma2A_FC, 
                              dfun = d_function, 
                              rfun = r_function,
                              sdtuning = 0.000022, #previous: 0.000045
                              betaA =  betaA_mat[it-1,], 
                              alphaA = alphaA_mat[it-1,],
                              D = D,
                              P = P,
                              A = A)
    sigma2A_vec[it] <- sigma2A_result
    
    
    
    betaD1_result <- MH_Update(x0 = betaD1_mat[it-1,], 
                               logFC = betaD1_FC, 
                               dfun = d_function, 
                               rfun = r_function,
                               sdtuning = 0.5, #previous: 0.6
                               alphaD = alphaD_mat[it-1,],
                               betaD2 = betaD2_mat[it-1,],
                               betaD3 = betaD3_mat[it-1,],
                               D = D,
                               P = P,
                               A = A)
    betaD1_mat[it,] <- betaD1_result
    
    
    
    betaD2_result <- MH_Update(x0 = betaD2_mat[it-1,], 
                               logFC = betaD2_FC, 
                               dfun = d_function, 
                               rfun = r_function,
                               sdtuning = 0.5, #previous: 0.45
                               betaD1 = betaD1_mat[it-1,],
                               betaD3 = betaD3_mat[it-1,],
                               alphaD = alphaD_mat[it-1,],
                               D = D,
                               P = P,
                               A = A)
    betaD2_mat[it,] <- betaD2_result
    
    
    
    betaD3_result <- MH_Update(x0 = betaD3_mat[it-1,], 
                               logFC = betaD3_FC, 
                               dfun = d_function, 
                               rfun = r_function,
                               sdtuning = 1.7, #previous: 1.7 
                               betaD1 = betaD1_mat[it-1,],
                               betaD2 = betaD2_mat[it-1,],
                               alphaD = alphaD_mat[it-1,],
                               D = D,
                               P = P,
                               A = A)
    betaD3_mat[it,] <- betaD3_result
    
    
    alphap_result <- MH_Update(x0 = alphap_mat[it-1,], 
                              logFC = alphap_FC, 
                              dfun = d_function, 
                              rfun = r_function,
                              sdtuning = 0.000008, #previous: 0.000035
                              sigma2p = sigma2p_vec[it-1], 
                              betaP = betaP_mat[it-1,],
                              D = D,
                              P = P,
                              A = A)
    alphap_mat[it,] = alphap_result
    
    
    alphaA_result <- MH_Update(x0 = alphaA_mat[it-1,], 
                              logFC = alphaA_FC, 
                              dfun = d_function, 
                              rfun = r_function,
                              sdtuning = 0.00003, #previous: 0.000035
                              sigma2A = sigma2A_vec[it-1], 
                              betaA =  betaA_mat[it-1,], 
                              D = D,
                              P = P,
                              A = A)
    alphaA_mat[it,] <- alphaA_result
    

    
    alphaD_result <- MH_Update(x0 = alphaD_mat[it-1,], 
                               logFC = alphaD_FC, 
                               dfun = d_function, 
                               rfun = r_function,
                               sdtuning = 0.6, #previous: 0.275
                               betaD1 = betaD1_mat[it-1,],
                               betaD2 = betaD2_mat[it-1,],
                               betaD3 = betaD3_mat[it-1,],
                               D = D,
                               P = P,
                               A = A)
    alphaD_mat[it,] <- alphaD_result
    
    
    D_result <- draw_D(D = D, 
                       missing_ind_D = missing_ind_D, 
                       propose_proportion = 0.05, # 0.05
                       betaP = betaP_mat[it-1,], 
                       alphap = alphap_mat[it-1,],
                       sigma2p = sigma2p_vec[it-1], 
                       betaD1 = betaD1_mat[it-1,], 
                       betaD2 = betaD2_mat[it-1,], 
                       betaD3 = betaD3_mat[it-1,], 
                       alphaD = alphaD_mat[it-1,],
                       P = P,
                       A = A)
    D_mat[it,] <- D_result[missing_ind_D]
    D[missing_ind_D] <- D_result[missing_ind_D]
    
    
    P_result <- draw_P(P = P, 
                     missing_ind_P = missing_ind_P,
                     propose_proportion = 0.30, # 0.15
                     sdtuning = 1e-06, #previous: 1e-07
                     betaP = betaP_mat[it-1,],
                     alphap = alphap_mat[it-1,],
                     sigma2p = sigma2p_vec[it-1],
                     D = D,
                     A = A)
    P_mat[it,] <- P_result[missing_ind_P]
    P[missing_ind_P] <- P_result[missing_ind_P]
    
    
    A_result <- draw_A(A = A, 
                    missing_ind_A = missing_ind_A,
                    propose_proportion = 0.30, # 0.15
                    sdtuning = 1e-06, #previous: 1e-07  
                    betaA =  betaA_mat[it-1,],
                    alphaA = alphaA_mat[it-1,],
                    sigma2A = sigma2A_vec[it-1],
                    D = D,
                    P = P)
    A_mat[it,] <- A_result[missing_ind_A]
    A[missing_ind_A] <- A_result[missing_ind_A]
    
    
    x <- c(betaP_mat[it,], sigma2p_vec[it], betaA_mat[it,], sigma2A_vec[it], 
            betaD1_mat[it,], betaD2_mat[it,], betaD3_mat[it,], alphap_mat[it,], 
            alphaA_mat[it,], alphaD_mat[it,], D_mat[it,], P_mat[it,], A_mat[it,])
    cat(paste0(x, collapse = ","), file = filename, append=TRUE)
    cat("\n", file = filename, append=TRUE) 
      
  }
}
```

```{r, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
#### Implementation of MH-Gibbs Algorithm =====================================

#time1 <- Sys.time() 
#MH_Gibbs(n_iter = 25000, filename = "chain1.csv")
#MH_Gibbs(n_iter = 25000, filename = "chain2.csv")
#MH_Gibbs(n_iter = 25000, filename = "chain3.csv")
#time2 <- Sys.time() 
#(runtime <- time2-time1)
```

\ 

### Summary of MCMC Results

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Analyzing Results
# Summary and Plots
getsummary <- function(data, BurPer = 0){
  
  vmean <- vlower <- vupper <- vsd <- NULL
  pos <- neg <- eq0 <- NULL
  
  for(i in 1:ncol(data)){
    param <- data[-(1:BurPer),i]
    vmean[i] <- mean(param)
    vlower[i] <- quantile(param, 0.025)
    vupper[i] <- quantile(param, 0.975)
    vsd[i] <- sd(param)
    pos[i] <- sum(1*(param > 0))/length(param)
    neg[i] <- sum(1*(param < 0))/length(param)
    eq0[i] <- sum(1*(param == 0))/length(param)
  }
    
  df <- data.frame(Parameter = colnames(data),
             Post.Mean = vmean,
             Lower95 = vlower,
             Upper95 = vupper,
             Post.SD = vsd,
             Accept_Rate = round(1 - rejectionRate(data),4),
             Pgreat0 = pos,
             Pless0 = neg,
             Pequal0 = eq0)
  rownames(df) <- NULL
  df
}

BurPer <- 5000
#results_list <- as.mcmc.list(lapply(1:3, 
#                                    function(x){as.mcmc(read.csv(paste0("chain", 
#                                                                        x, 
#                                                                        ".csv"),
#                                                                 skip = BurPer))}))
#combined_list <- combine.MCMC(results_list)

c1 <- as.mcmc(read.csv(paste0("chain1.csv")))
c2 <- as.mcmc(read.csv(paste0("chain2.csv")))
c3 <- as.mcmc(read.csv(paste0("chain3.csv")))

x1 <- window(c1,start = BurPer+1,dim(c1)[1])
x2 <- window(c2,start = BurPer+1,dim(c1)[1])
x3 <- window(c3,start = BurPer+1,dim(c1)[1])

results_list <- as.mcmc.list(list(x1,x2,x3))
combined_list <- combine.MCMC(results_list)
options(digits = 4)
(mcmc_summary = getsummary(data = combined_list, BurPer = BurPer))
```
 
\ 

### Trace Plots and Convergence Diagnostic {.tabset}

#### Hide

\ 

#### Gelman Rubin Diagnostics

```{r, echo=FALSE, message=FALSE, warning=FALSE}
gelman.diag(results_list, multivariate = FALSE)
```
\

#### Trace Plots

```{r, echo=FALSE, message=FALSE, warning=FALSE}
par(mfrow=c(2,2))
traceplot(results_list)
```

\ 

## Predictions {.tabset}

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Use examples from data set, to make predictions
# Most important IDs: 761 (will be predicting missing data), 769, 772, and 775

#### DATA ===========================================================
dog_ids <- leish$ID 

dt_P <- as.matrix(leish[,c("PCR_Load_TP1","PCR_Load_TP2", 
                           "PCR_Load_TP3", "PCR_Load_TP4",
                           "PCR_Load_TP5", "PCR_Load_TP6",
                           "PCR_Load_TP7")])
dt_P[missP] <- ifelse(mcmc_summary$Post.Mean[118:137] < 0, 0,
                      mcmc_summary$Post.Mean[118:137]*generalmaxP)


dt_D <- as.matrix(leish[,c("DStatus_TP1","DStatus_TP2", 
                           "DStatus_TP3", "DStatus_TP4",
                           "DStatus_TP5", "DStatus_TP6",
                           "DStatus_TP7")])
dt_D[missD] <- round(mcmc_summary$Post.Mean[45:117])

dt_A <- as.matrix(leish[,c("SLA.ELISA.OD.ratio_TP1",
                           "SLA.ELISA.OD.ratio_TP2", 
                           "SLA.ELISA.OD.ratio_TP3", 
                           "SLA.ELISA.OD.ratio_TP4",
                           "SLA.ELISA.OD.ratio_TP5",
                           "SLA.ELISA.OD.ratio_TP6",
                           "SLA.ELISA.OD.ratio_TP7")])
dt_A[missA] <- ifelse(mcmc_summary$Post.Mean[118:137] < 0, 0,
                      mcmc_summary$Post.Mean[118:137]*generalmaxA)

dt_X <- cbind(1, leish$Age_group, 
              leish$Ever.SNAP.Positive,
              leish$DPP_Enroll/generalmaxDPP,
              leish$treatment_group)

#### MCMC Results ==========================================================

# Posteriors
results <- combine.mcmc(results_list)

#### Posterior Predictive Distributions =======================================

nsim = 1000
time_pred = 7

P_s1 = P_s2 = P_s3 = P_s4 = matrix(data = NA, nrow = nsim, ncol = time_pred)
A_s1 = A_s2 = A_s3 = A_s4 = matrix(data = NA, nrow = nsim, ncol = time_pred)
D_s1 = D_s2 = D_s3 = D_s4 = matrix(data = NA, nrow = nsim, ncol = time_pred)

P_s1[,1:2] <- cbind(rep(dt_P[which(dog_ids =="761"),1], nsim)/generalmaxP, 
                   rep(dt_P[which(dog_ids =="761"),2], nsim)/generalmaxP)
P_s2[,1:2] <- cbind(rep(dt_P[which(dog_ids =="772"),1], nsim)/generalmaxP,
                  rep(dt_P[which(dog_ids =="772"),2], nsim)/generalmaxP)
P_s3[,1:2] <- cbind(rep(dt_P[which(dog_ids =="769"),1], nsim)/generalmaxP,
                  rep(dt_P[which(dog_ids =="769"),2], nsim)/generalmaxP)
P_s4[,1:2] <- cbind(rep(dt_P[which(dog_ids =="775"),1], nsim)/generalmaxP,
                  rep(dt_P[which(dog_ids =="775"),2], nsim)/generalmaxP)

A_s1[,1:2] <- cbind(rep(dt_A[which(dog_ids =="761"),1], nsim)/generalmaxA, 
                   rep(dt_A[which(dog_ids =="761"),2], nsim)/generalmaxA)
A_s2[,1:2] <- cbind(rep(dt_A[which(dog_ids =="772"),1], nsim)/generalmaxA,
                  rep(dt_A[which(dog_ids =="772"),2], nsim)/generalmaxA)
A_s3[,1:2] <- cbind(rep(dt_A[which(dog_ids =="769"),1], nsim)/generalmaxA,
                  rep(dt_A[which(dog_ids =="769"),2], nsim)/generalmaxA)
A_s4[,1:2] <- cbind(rep(dt_A[which(dog_ids =="775"),1], nsim)/generalmaxA,
                  rep(dt_A[which(dog_ids =="775"),2], nsim)/generalmaxA)

D_s1[,1:2] <- cbind(rep(dt_D[which(dog_ids =="761"),1], nsim), 
                   rep(dt_D[which(dog_ids =="761"),2], nsim))
D_s2[,1:2] <- cbind(rep(dt_D[which(dog_ids =="772"),1], nsim),
                  rep(dt_D[which(dog_ids =="772"),2], nsim))
D_s3[,1:2] <- cbind(rep(dt_D[which(dog_ids =="769"),1], nsim),
                  rep(dt_D[which(dog_ids =="769"),2], nsim))
D_s4[,1:2] <- cbind(rep(dt_D[which(dog_ids =="775"),1], nsim),
                  rep(dt_D[which(dog_ids =="775"),2], nsim))

X = dt_X[c(which(dog_ids =="761"), 
           which(dog_ids =="772"),
           which(dog_ids =="769"),
           which(dog_ids =="775")),]

# predict
for(t in 3:time_pred){

    for(s in 1:nsim){
      random_param <- apply(results, 2, function(x){sample(x,1)})
      
      Data <- data.frame(P = c(P_s1[s,t-1],P_s2[s,t-1],P_s3[s,t-1],P_s4[s,t-1]),
                         A = c(A_s1[s,t-1],A_s2[s,t-1],A_s3[s,t-1],A_s4[s,t-1]),
                         D1 = ifelse(c(D_s1[s,t-1],
                                       D_s2[s,t-1],
                                       D_s3[s,t-1],
                                       D_s4[s,t-1]) == 1,1,0),
                         D2 = ifelse(c(D_s1[s,t-1],
                                       D_s2[s,t-1],
                                       D_s3[s,t-1],
                                       D_s4[s,t-1]) == 2,1,0),
                         D3 = ifelse(c(D_s1[s,t-1],
                                       D_s2[s,t-1],
                                       D_s3[s,t-1],
                                       D_s4[s,t-1]) == 3,1,0),
                         D1diffP = c(ifelse(D_s1[s,t-1] == 1,1,0)*(P_s1[s,t-1]-P_s1[s,t-2]), 
                                     ifelse(D_s2[s,t-1]== 1,1,0)*(P_s2[s,t-1]-P_s2[s,t-2]),
                                     ifelse(D_s3[s,t-1]== 1,1,0)*(P_s3[s,t-1]-P_s3[s,t-2]),
                                     ifelse(D_s4[s,t-1]== 1,1,0)*(P_s4[s,t-1]-P_s4[s,t-2])),
                         D2diffP = c(ifelse(D_s1[s,t-1] == 2,1,0)*(P_s1[s,t-1]-P_s1[s,t-2]), 
                                     ifelse(D_s2[s,t-1]== 2,1,0)*(P_s2[s,t-1]-P_s2[s,t-2]),
                                     ifelse(D_s3[s,t-1]== 2,1,0)*(P_s3[s,t-1]-P_s3[s,t-2]),
                                     ifelse(D_s4[s,t-1]== 2,1,0)*(P_s4[s,t-1]-P_s4[s,t-2])),
                         D3diffP = c(ifelse(D_s1[s,t-1] == 3,1,0)*(P_s1[s,t-1]-P_s1[s,t-2]), 
                                     ifelse(D_s2[s,t-1]== 3,1,0)*(P_s2[s,t-1]-P_s2[s,t-2]),
                                     ifelse(D_s3[s,t-1]== 3,1,0)*(P_s3[s,t-1]-P_s3[s,t-2]),
                                     ifelse(D_s4[s,t-1]== 3,1,0)*(P_s4[s,t-1]-P_s4[s,t-2])),
                         D1diffA = c(ifelse(D_s1[s,t-1] == 1,1,0)*(A_s1[s,t-1]-A_s1[s,t-2]), 
                                     ifelse(D_s2[s,t-1]== 1,1,0)*(A_s2[s,t-1]-A_s2[s,t-2]),
                                     ifelse(D_s3[s,t-1]== 1,1,0)*(A_s3[s,t-1]-A_s3[s,t-2]),
                                     ifelse(D_s4[s,t-1]== 1,1,0)*(A_s4[s,t-1]-A_s4[s,t-2])),
                         D2diffA = c(ifelse(D_s1[s,t-1] == 2,1,0)*(A_s1[s,t-1]-A_s1[s,t-2]), 
                                     ifelse(D_s2[s,t-1]== 2,1,0)*(A_s2[s,t-1]-A_s2[s,t-2]),
                                     ifelse(D_s3[s,t-1]== 2,1,0)*(A_s3[s,t-1]-A_s3[s,t-2]),
                                     ifelse(D_s4[s,t-1]== 2,1,0)*(A_s4[s,t-1]-A_s4[s,t-2])),
                         D3diffA = c(ifelse(D_s1[s,t-1] == 3,1,0)*(A_s1[s,t-1]-A_s1[s,t-2]), 
                                     ifelse(D_s2[s,t-1]== 3,1,0)*(A_s2[s,t-1]-A_s2[s,t-2]),
                                     ifelse(D_s3[s,t-1]== 3,1,0)*(A_s3[s,t-1]-A_s3[s,t-2]),
                                     ifelse(D_s4[s,t-1]== 3,1,0)*(A_s4[s,t-1]-A_s4[s,t-2])))
      Data$subject <- rep(1:4, each = 1)  
      
      # Create model matrices
      X1 <- with(Data, cbind(P, D1*P, D2*P, D3*P, D1diffP, D2diffP, D3diffP, 
                             D1diffA, D2diffA, D3diffA))
      X2 <- with(Data, cbind(P, A, D1diffP, D2diffP, D3diffP, 
                             D1diffA, D2diffA, D3diffA))
      X3 <- cbind(Data$D1, Data$D1*Data$P, Data$D1*Data$P)
      X4 <- cbind(Data$D2, Data$D2*Data$P, Data$D2*Data$P)
      X5 <- cbind(Data$D3, Data$D3*Data$P, Data$D3*Data$P)
      
      M <- model.matrix(~ as.factor(Data$subject) - 1)
      idx <- t(apply(M, 2, function(X){c(min(which(X == 1)), 
                                         max(which(X == 1)))}))
      idx <- data.frame(start_idx = idx[,1], stop_idx = idx[,2])
      
      # pathogen load
      pmean = (X1[idx[,2],] %*% as.numeric(random_param[1:10])) +
        (X[idx[,2],] %*% as.numeric(random_param[30:34])) 
      pvar = as.numeric(random_param[11])
      P_s1[s,t] <- rnorm(1, mean = pmean[1,1], sd = sqrt(pvar))
      P_s2[s,t] <- rnorm(1, mean = pmean[2,1], sd = sqrt(pvar))
      P_s3[s,t] <- rnorm(1, mean = pmean[3,1], sd = sqrt(pvar))
      P_s4[s,t] <- rnorm(1, mean = pmean[4,1], sd = sqrt(pvar))
      
      # antibodies
      amean = (X2[idx[,2],] %*% as.numeric(random_param[12:19])) +
        (X[idx[,2],] %*% as.numeric(random_param[35:39])) 
      avar = as.numeric(random_param[20])
      A_s1[s,t] <- rnorm(1, mean = amean[1,1], sd = sqrt(avar))
      A_s2[s,t] <- rnorm(1, mean = amean[2,1], sd = sqrt(avar))
      A_s3[s,t] <- rnorm(1, mean = amean[3,1], sd = sqrt(avar))
      A_s4[s,t] <- rnorm(1, mean = amean[4,1], sd = sqrt(avar))
      
      # predicted prob and disease status
      num1 = exp((X3[idx[,2],] %*% as.numeric(random_param[21:23])) +
                   (X[idx[,2],] %*% as.numeric(random_param[30:34])))
      
      num2 = exp((X4[idx[,2],] %*% as.numeric(random_param[24:26])) +
                   (X[idx[,2],] %*% as.numeric(random_param[30:34])))
      
      num3 = exp((X5[idx[,2],] %*% as.numeric(random_param[27:29])) +
                   (X[idx[,2],] %*% as.numeric(random_param[30:34])))
      
      den = 1 + num1 + num2 + num3
      
      pi1 <- num1/den
      pi2 <- num2/den
      pi3 <- num3/den 
      pi4 <- 1 - num1/den - num2/den - num3/den
      
      
      D_s1[s,t] <- which.max(rmultinom(1,1,prob = c(pi1[1,1],
                                                    pi2[1,1],
                                                    pi3[1,1],
                                                    pi4[1,1])))
                      
      D_s2[s,t] <- which.max(rmultinom(1,1,prob = c(pi1[2,1],
                                                    pi2[2,1],
                                                    pi3[2,1],
                                                    pi4[2,1])))
      
      D_s3[s,t] <- which.max(rmultinom(1,1,prob = c(pi1[3,1],
                                                    pi2[3,1],
                                                    pi3[3,1],
                                                    pi4[3,1])))
      
      D_s4[s,t] <- which.max(rmultinom(1,1,prob = c(pi1[4,1],
                                                    pi2[4,1],
                                                    pi3[4,1],
                                                    pi4[4,1])))
      
    }
}

#### Plot =============================================================


pred.plot <- function(subj = 1, 
                      rescaled = TRUE,
                      pred_time = 7){
  
  if(subj == 1){
    pathogen = P_s1
    antibodies = A_s1
    Dstatus = D_s1
  } else if (subj == 2){
    pathogen = P_s2
    antibodies = A_s2
    Dstatus = D_s2
  } else if(subj == 3){
    pathogen = P_s3
    antibodies = A_s3
    Dstatus = D_s3
  } else if(subj == 4){
    pathogen = P_s4
    antibodies = A_s4
    Dstatus = D_s4
  } else {cat("Error: No valid subject")}
  
  
  if(rescaled == TRUE){
    pathogen = pathogen*generalmaxP
    antibodies = antibodies*generalmaxA
  } 
  
  if(subj == 1) {
    id = 761
  } else if(subj == 2){
    id = 772
  } else if(subj == 3){
    id = 769
  } else if(subj == 4){
    id = 775
  } else {cat("Error: No valid subject")}
  
  pathogen <- ifelse(pathogen < 0, 0, pathogen)
  antibodies <- ifelse(antibodies < 0, 0, antibodies)
  
  lowerp80 = apply(pathogen, 2, function(x) quantile(x,0.1))
  upperp80 = apply(pathogen, 2, function(x) quantile(x,0.9))
  
  lowera80 = apply(antibodies, 2, function(x) quantile(x,0.1))
  uppera80 = apply(antibodies, 2, function(x) quantile(x,0.9))
  
  lowerd80 = apply(Dstatus, 2, function(x) quantile(x,0.1))
  upperd80 = apply(Dstatus, 2, function(x) quantile(x,0.9))
  
  
  lowerp95 = apply(pathogen, 2, function(x) quantile(x,0.025))
  upperp95 = apply(pathogen, 2, function(x) quantile(x,0.975))
  
  lowera95 = apply(antibodies, 2, function(x) quantile(x,0.025))
  uppera95 = apply(antibodies, 2, function(x) quantile(x,0.975))
  
  lowerd95 = apply(Dstatus, 2, function(x) quantile(x,0.025))
  upperd95 = apply(Dstatus, 2, function(x) quantile(x,0.975))
  
  
  par(mfrow=c(1,1))
  plot(1:pred_time, pathogen[1,], 
       xlab = "Time Point",
       ylab = "Pathogen Load",
       main = paste0("Predicted Pathogen Load for Subject ", subj),
       col = "darkgray",
       type = "l",
       ylim = c(0,100000),
       panel.first = ggbg("gray92"))
  for(i in 2:nrow(pathogen)){
      lines(1:pred_time, pathogen[i,], col = "darkgray")
  }
  lines(1:pred_time, lowerp80, col = "black", lwd = 3, lty = 2)
  lines(1:pred_time, upperp80, col = "black", lwd = 3, lty = 2)
  lines(1:pred_time, lowerp95, col = "royalblue3", lwd = 3, lty = 2)
  lines(1:pred_time, upperp95, col = "royalblue3", lwd = 3, lty = 2)
  lines(1:pred_time, dt_P[which(dog_ids==id),], col = "maroon",lwd = 3, 
        lty=1, type="b")
  lines(1:pred_time, 
        apply(pathogen,2,function(x)mean(x)),
        col = "red",
        lwd = 3)
  grid(NULL, NULL, col = "white")
  legend("topleft", legend=c("Mean", "Observed", 
                             "80% CI Bounds", "95% CI Bounds"), 
       lty = c(1,1,2,2), lwd = c(3,3,3,3), col = c("red", "maroon", 
                                                   "black", "royalblue3"),
       inset = 0.02, cex = 0.85, seg.len=3.5, pch = c(NA,19,NA,NA))
  

  plot(1:pred_time, antibodies[1,], 
       xlab = "Time Point",
       ylab = "Antibody Levels",
       main = paste0("Predicted Antibody Levels for Subject ", subj),
       col = "darkgray",
       type = "l",
       ylim = c(0,25),
       panel.first = ggbg("gray92"))
  for(i in 2:nrow(antibodies)){
      lines(1:pred_time, antibodies[i,], col = "darkgray")
  }
  lines(1:pred_time, lowera80, col = "black", lwd = 3, lty = 2)
  lines(1:pred_time, uppera80, col = "black", lwd = 3, lty = 2)
  lines(1:pred_time, lowera95, col = "royalblue3", lwd = 3, lty = 2)
  lines(1:pred_time, uppera95, col = "royalblue3", lwd = 3, lty = 2)
  lines(1:pred_time, dt_A[which(dog_ids==id),], col = "forestgreen",lwd = 3, 
        lty=1, type="b")
  lines(1:pred_time, 
        apply(antibodies,2,function(x)mean(x)),
        col = "darkgreen",
        lwd = 3)
  grid(NULL, NULL, col = "white")
  legend("topleft", legend=c("Mean", "Observed", 
                             "80% CI Bounds", "95% CI Bounds"), 
       lty = c(1,1,2,2), lwd = c(3,3,3,3), col = c("darkgreen", "forestgreen", 
                                                   "black", "royalblue3"),
       inset = 0.02, cex = 0.85, seg.len=3.5, pch = c(NA,19,NA,NA))
  

  
  df <- matrix(data = 0, 4, pred_time)
  if(Dstatus[1,1]==1) {
    df[,1] = c(1,0,0,0)
    } else if(Dstatus[1,1]==2) {
      df[,1] = c(0,1,0,0)
      } else if(Dstatus[1,1]==3) {
          df[,1] = c(0,0,1,0)
          } else {df[,1] = c(0,0,0,1)}
  
  if(Dstatus[2,1]==1) {
    df[,2] = c(1,0,0,0)
    } else if(Dstatus[2,1]==2) {
      df[,2] = c(0,1,0,0)
      } else if(Dstatus[2,1]==3) {
          df[,2] = c(0,0,1,0)
          } else {df[,2] = c(0,0,0,1)}
  
  for(i in 3:pred_time){
    df[,i] <- table(Dstatus[,i])/nsim
  }
  
  colnames(df) <- paste0("TP", 1:7)
  rownames(df) <- paste0("DS = ", 1:4)
 
  kable(df, "html", booktabs = T,
        caption = "Marginal probabilities of disease status (DS) at each time point") %>%
  kable_styling(latex_options = c("striped"), position = "center",
                full_width = F)
  
  plot(1:pred_time, df[1,], 
       xlab = "Time Point",
       ylab = "Probability",
       main = paste0("Marginal probabilities of disease status (D) for Subject ", subj),
       col = "cadetblue",
       type = "b",
       ylim = c(0,1),
       panel.first = ggbg("gray92"),
       lwd = 2, lty = 1)
  grid(NULL, NULL, col = "white")
  lines(1:pred_time, df[1,], col = "cadetblue", lwd = 2.5, lty = 1, type = "b", pch = 18)
  lines(1:pred_time, df[2,], col = "cornflowerblue", lwd = 2.5, lty = 1, type = "b", pch = 18)
  lines(1:pred_time, df[3,], col = "chocolate", lwd = 2.5, lty = 1, type = "b", pch = 17)
  lines(1:pred_time, df[4,], col = "darkred", lwd = 2.5, lty = 1, type = "b", pch = 16)
  legend(0.95,0.95, legend=c("D=1","D=2","D=3","D=4"), 
       lty = c(1,1,1,1), lwd = 2.5, col = c("cadetblue", "cornflowerblue", 
                                          "chocolate", "darkred"),
       inset = 0.02, cex = 0.9, seg.len=2, pch = c(19:16))
  
}
```

\ 

Data (4 subjects within dataset):
```{r, echo=FALSE, message=FALSE, warning=FALSE}
pat = c(P_s1[1,1], P_s2[1,1], P_s3[1,1], P_s4[1,1], 
        P_s1[1,2], P_s2[1,2], P_s3[1,2], P_s4[1,2])
ant = c(A_s1[1,1], A_s2[1,1], A_s3[1,1], A_s4[1,1], 
        A_s1[1,2], A_s2[1,2], A_s3[1,2], A_s4[1,2])
dis = c(D_s1[1,1], D_s2[1,1], D_s3[1,1], D_s4[1,1],
        D_s1[1,2], D_s2[1,2], D_s3[1,2], D_s4[1,2])
xdat = X
df1 = cbind(data.frame(rep(1:2,each=4), rep(1:4,2),pat,ant,dis),xdat)
colnames(df1) <- c("Time Point", "New Subject", "Pathogen Load", "Antibodies", 
                  "Disease Status", "Intercept", "Age Group", "Ever Snap Positive", 
                  "DPP", "Treatment Group")
df1
```

\ 

### Hide

\ 

### Subject 1
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.asp=1.3}
pred.plot(subj = 1)
```

\ 

### Subject 2

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.asp=1.3}
pred.plot(subj = 2)
```

\ 

### Subject 3

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.asp=1.3}
pred.plot(subj = 3)
```

\ 

### Subject 4

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.asp=1.3}
pred.plot(subj = 4)
```



\ 

\ 

\ 




















