---
title: "Supplementary Document: MCMC Summary"
subtitle: "Within-Host Bayesian Joint Modeling of Longitudinal and Time-to-Event Data of Leishmania Infection"
author: "Felix M. Pabon-Rodriguez, Grant D. Brown, Breanna M. Scorza, and Christine A. Petersen"
date: ''
output: pdf_document
---
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Loading libraries
options(repos="https://cran.rstudio.com")
install_load <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg))
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)
}

my_packages <- c("readxl", "dplyr", "coda", "ggplot2",
                 "lattice", "bayesplot", "BayesPostEst", "ggmcmc", 
                 "RCurl", "truncnorm", "kableExtra", "mvtnorm", "rlist", 
                 "extraDistr", "msm", "tmvtnorm", "runjags", "plotrix",
                 "lubridate", "ggpubr", "stringr", "nimble",
                 "igraph", "parallel", "doParallel", "MCMCvis")
invisible(install_load(my_packages))


# Results
results_file <- readRDS(file = "resultsBJLS_model.rds")
results_mcmc <- as.mcmc.list(lapply(1:3, function(x){as.mcmc(results_file[[x]])}))
combined_list <- combine.MCMC(results_mcmc)
options(digits = 4)
```

```{r, echo=FALSE}

getsummary <- function(data, BurPer = 0){
  
  vmean <- vlower <- vupper <- vsd <- NULL
  pos <- neg <- vmedian <- strong <- NULL
  
  for(i in 1:ncol(data)){
    param <- data[-(1:BurPer),i]
    vmean[i] <- mean(param)
    vmedian[i] <- median(param)
    vlower[i] <- quantile(param, 0.025)
    vupper[i] <- quantile(param, 0.975)
    vsd[i] <- sd(param) 
    pos[i] <- sum(1*(param > 0))/length(param)
    neg[i] <- sum(1*(param < 0))/length(param)
    strong[i] <- ifelse((pos[i] > 0.65)||(neg[i] > 0.65), "*", "")
  }
    
  df <- data.frame(Parameter = colnames(data),
             Post.Mean = vmean,
             Post.Median = vmedian,
             Post.SD = vsd,
             Lower95 = vlower,
             Upper95 = vupper,
             P.great.0 = pos,
             P.less.0 = neg,
             Evidence = strong)
  rownames(df) <- NULL
  df
}

```

**Supplementary Table.** Summary of MCMC results for all interpretable parameters in the model. Each column represents the following: (1) Posterior mean, (2) Posterior median, (3) Posterior standard deviation (SD), (4) Lower bound of 95% credible interval (Cr-I), (5) Upper bound of 95% Cr-I, (6) Posterior probability of parameter being positive, (7) Posterior probability of parameter being negative, and (8) Indicator for strength of evidence based on posterior probability. The symbol \* in last column is used to indicate which parameter shows a posterior probability greater or equal than 0.65. The results in this table are separated by model components. Last part of the table summarizes parameters associated with the survival submodel.

\vspace{0.5cm}

**Pathogen Load (P)**

```{r, echo=FALSE}

clnamesP <- c(colnames(combined_list[,grepl("betaP",colnames(combined_list))]),
             colnames(combined_list[,grepl("alphaP",colnames(combined_list))]))

dfP <- getsummary(data = combined_list[,clnamesP])

kbl(dfP, booktabs = T, longtable = T) %>%
kable_styling(latex_options = "striped")
```

\newpage

**Antibody Levels (A)**

```{r, echo=FALSE}

clnamesA <- c(colnames(combined_list[,grepl("betaA",colnames(combined_list))]),
             colnames(combined_list[,grepl("alphaA",colnames(combined_list))]))

dfA <- getsummary(data = combined_list[,clnamesA])

kbl(dfA, booktabs = T, longtable = T) %>%
kable_styling(latex_options = "striped")
```

\newpage

**Disease Status (D)**

```{r, echo=FALSE}

clnamesD <- c(colnames(combined_list[,grepl("betaD2",colnames(combined_list))]),
             colnames(combined_list[,grepl("betaD3",colnames(combined_list))]),
             colnames(combined_list[,grepl("alphaD2",colnames(combined_list))]),
             colnames(combined_list[,grepl("alphaD3",colnames(combined_list))]))

dfD <- getsummary(data = combined_list[,clnamesD])

kbl(dfD, booktabs = T, longtable = T) %>%
kable_styling(latex_options = "striped")
```

\newpage

**Inflammatory Responses (I1, I2, I3)**

```{r, echo=FALSE}

clnamesI <- c(colnames(combined_list[,grepl("betaI1",colnames(combined_list))]),
             colnames(combined_list[,grepl("betaI2",colnames(combined_list))]),
             colnames(combined_list[,grepl("betaI3",colnames(combined_list))]),
             colnames(combined_list[,grepl("alphaI1",colnames(combined_list))]),
             colnames(combined_list[,grepl("alphaI2",colnames(combined_list))]),
             colnames(combined_list[,grepl("alphaI3",colnames(combined_list))]))

dfI <- getsummary(data = combined_list[,clnamesI])

kbl(dfI, booktabs = T, longtable = T) %>%
kable_styling(latex_options = "striped")
```

\newpage

**Regulatory Responses (R1, R2, R3)**

```{r, echo=FALSE}

clnamesR <- c(colnames(combined_list[,grepl("betaR1",colnames(combined_list))]),
             colnames(combined_list[,grepl("betaR2",colnames(combined_list))]),
             colnames(combined_list[,grepl("betaR3",colnames(combined_list))]),
             colnames(combined_list[,grepl("alphaR1",colnames(combined_list))]),
             colnames(combined_list[,grepl("alphaR2",colnames(combined_list))]),
             colnames(combined_list[,grepl("alphaR3",colnames(combined_list))]))

dfR <- getsummary(data = combined_list[,clnamesR])

kbl(dfR, booktabs = T, longtable = T) %>%
kable_styling(latex_options = "striped")
```

\newpage

**Standard Deviations and Covariance Matrix**

```{r, echo=FALSE}

clnamesSD <- c("sigmaP", "sigmaA",
              colnames(combined_list[,grepl("SigmaIR",colnames(combined_list))]))

dfSD <- getsummary(data = combined_list[,clnamesSD])

kbl(dfSD, booktabs = T, longtable = T) %>%
kable_styling(latex_options = "striped")
```

\newpage

**Hazard Parameters**

```{r, echo=FALSE}

clnamesSurv <- c("shape", "scale",
              colnames(combined_list[,grepl("assoc",colnames(combined_list))]),
              colnames(combined_list[,grepl("gammaS",colnames(combined_list))]))

dfSurv <- getsummary(data = combined_list[,clnamesSurv])

kbl(dfSurv, booktabs = T, longtable = T) %>%
kable_styling(latex_options = "striped")
```
